### 2019-08-19 Tables/te.tex
Initial git commit

% Table generated by Excel2LaTeX from sheet 'te'
\begin{tabular}{rrrrrrrrrrr}
\toprule
      & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} \\
\midrule
\midrule
\multicolumn{1}{l}{} & \multicolumn{2}{c}{(1)} & \multicolumn{2}{c}{(2)} & \multicolumn{2}{c}{(3)} & \multicolumn{2}{c}{(4)} & \multicolumn{2}{c}{(5)} \\
\midrule
\midrule
\multicolumn{1}{l}{No choice/Fee} & \multicolumn{1}{l}{0.11***} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{-5.88*} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{0.076***} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{0.068*} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{0.11***} & \multicolumn{1}{l}{} \\
\multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.023)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(3.20)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.026)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.038)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.015)} & \multicolumn{1}{l}{} \\
\multicolumn{1}{l}{No choice/No fee} & \multicolumn{1}{l}{-0.0048} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{-0.30} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{-0.011} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{0.092*} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{0.053***} & \multicolumn{1}{l}{} \\
\multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.022)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(3.71)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.024)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.051)} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.014)} & \multicolumn{1}{l}{} \\
      &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} 0.14***} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} -6.81*} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} 0.097**} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} 0.41***} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} 0.29***} \\
\multicolumn{1}{l}{Choice/Fee} & \multicolumn{1}{l}{-0.0020} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.037)} & \multicolumn{1}{l}{1.58} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (3.82)} & \multicolumn{1}{l}{0.0044} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.042)} & \multicolumn{1}{l}{0.10**} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.096)} & \multicolumn{1}{l}{0.029**} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.034)} \\
\cmidrule{3-3}\cmidrule{5-5}\cmidrule{7-7}\cmidrule{9-9}\cmidrule{11-11}\multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.021)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} -0.018} & \multicolumn{1}{l}{(3.32)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} 2.89} & \multicolumn{1}{l}{(0.022)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} -0.0060} & \multicolumn{1}{l}{(0.039)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} 0.065*} & \multicolumn{1}{l}{(0.013)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} -0.000012} \\
      &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.021)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (3.49)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.022)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.039)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.012)} \\
\cmidrule{9-9}\cmidrule{11-11}      &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} -0.037} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} -8.66**} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} -0.058**} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} 0.073} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} 0.082***} \\
\multicolumn{1}{l}{Choice/No fee} & \multicolumn{1}{l}{-0.0066} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.026)} & \multicolumn{1}{l}{-3.31} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (4.09)} & \multicolumn{1}{l}{-0.018} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.029)} & \multicolumn{1}{l}{-0.0012} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.057)} & \multicolumn{1}{l}{0.019} & \multicolumn{1}{l}{\cellcolor[rgb]{ .929,  .929,  .929} (0.018)} \\
\cmidrule{3-3}\cmidrule{5-5}\cmidrule{9-9}\cmidrule{11-11}\multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.022)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} 0.0070} & \multicolumn{1}{l}{(3.43)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} -1.13} & \multicolumn{1}{l}{(0.023)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} -0.00055} & \multicolumn{1}{l}{(0.037)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} -0.035} & \multicolumn{1}{l}{(0.011)} & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} -0.010} \\
      &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.025)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (3.88)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.026)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.038)} &       & \multicolumn{1}{l}{\cellcolor[rgb]{ .859,  .859,  .859} (0.012)} \\
      &       &       &       &       &       &       &       &       &       &  \\
\multicolumn{1}{l}{Constant (Control)} & \multicolumn{1}{l}{0.44***} & \multicolumn{1}{l}{0.44***} & \multicolumn{1}{l}{92.4***} & \multicolumn{1}{l}{92.4***} & \multicolumn{1}{l}{-0.35***} & \multicolumn{1}{l}{-0.35***} & \multicolumn{1}{l}{0.99***} & \multicolumn{1}{l}{0.99***} & \multicolumn{1}{l}{0.12***} & \multicolumn{1}{l}{0.12***} \\
\multicolumn{1}{l}{} & \multicolumn{1}{l}{(0.015)} & \multicolumn{1}{l}{(0.015)} & \multicolumn{1}{l}{(2.53)} & \multicolumn{1}{l}{(2.53)} & \multicolumn{1}{l}{(0.017)} & \multicolumn{1}{l}{(0.017)} & \multicolumn{1}{l}{(0.028)} & \multicolumn{1}{l}{(0.028)} & \multicolumn{1}{l}{(0.0085)} & \multicolumn{1}{l}{(0.0085)} \\
      &       &       &       &       &       &       &       &       &       &  \\
\midrule
\multicolumn{1}{l}{Observations} & \multicolumn{1}{l}{13446} & \multicolumn{1}{l}{13446} & \multicolumn{1}{l}{6119} & \multicolumn{1}{l}{6119} & \multicolumn{1}{l}{13446} & \multicolumn{1}{l}{13446} & \multicolumn{1}{l}{13446} & \multicolumn{1}{l}{13446} & \multicolumn{1}{l}{13446} & \multicolumn{1}{l}{13446} \\
\multicolumn{1}{l}{R-sq} & \multicolumn{1}{l}{0.007} & \multicolumn{1}{l}{0.010} & \multicolumn{1}{l}{0.003} & \multicolumn{1}{l}{0.005} & \multicolumn{1}{l}{0.003} & \multicolumn{1}{l}{0.005} & \multicolumn{1}{l}{0.002} & \multicolumn{1}{l}{0.006} & \multicolumn{1}{l}{0.010} & \multicolumn{1}{l}{0.028} \\
\multicolumn{1}{l}{Dependent Var Mean} & \multicolumn{2}{c}{0.46} & \multicolumn{2}{c}{90.8} & \multicolumn{2}{c}{-0.34} & \multicolumn{2}{c}{1.04} & \multicolumn{2}{c}{0.16} \\
\midrule
\midrule
\rowcolor[rgb]{ .929,  .929,  .929} Status-quo & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  \\
\rowcolor[rgb]{ .859,  .859,  .859} Non status-quo & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  & \cellcolor[rgb]{ 1,  1,  1}  \\
\end{tabular}%