### 2020-01-20 Tables/te_pro_2.tex
Work in paper in winter break MX

% Table generated by Excel2LaTeX from sheet 'te_pro_2'
\begin{tabular}{lcccc|c}
\toprule
      & Paid loan & FC admin  & FC  admin disc & \multicolumn{1}{c}{FC subjective} & Number of payment \\
\midrule
\midrule
      & (1)   & (2)   & (3)   & \multicolumn{1}{c}{(4)} & (5) \\
\midrule
\midrule
No Choice/Fee & 0.11*** & -350.2*** & -339.7*** & \multicolumn{1}{c}{-791.1*} & 0.73*** \\
      & (0.022) & (121.7) & (111.2) & \multicolumn{1}{c}{(469.5)} & (0.19) \\
Constant & 0.42*** & 4071.9*** & 3612.2*** & \multicolumn{1}{c}{5449.2***} & 2.69*** \\
      & (0.052) & (309.9) & (278.9) & \multicolumn{1}{c}{(1219.6)} & (0.44) \\
      &       &       &       & \multicolumn{1}{c}{} &  \\
\midrule
Observations & 5085  & 5034  & 5034  & \multicolumn{1}{c}{3954} & 5085 \\
R-sq  & 0.026 & 0.017 & 0.015 & \multicolumn{1}{c}{0.024} & 0.010 \\
Control Mean & 0.44  & 3231.8 & 2915.7 & \multicolumn{1}{c}{4355.7} & 2.08 \\
\midrule
DOW FE & YES   & YES   & YES   & \multicolumn{1}{c}{YES} & YES \\
Branch FE & YES   & YES   & YES   & \multicolumn{1}{c}{YES} & YES \\
Number of pawns FE & YES   & YES   & YES   & \multicolumn{1}{c}{YES} & YES \\
Number of arms FE & YES   & YES   & YES   & \multicolumn{1}{c}{YES} & YES \\
\midrule
\midrule
      &       &       &       & \multicolumn{1}{c}{} &  \\
\midrule
      & \% of payment & Size of payment & Size of payment (disc) & Days to un-pledge & Reincidence \\
\midrule
\midrule
      & (6)   & (7)   & (8)   & (9)   & (10) \\
\midrule
\midrule
No Choice/Fee & 0.081*** & 34.9  & 47.1  & -8.51*** & 0.053*** \\
      & (0.025) & (50.9) & (42.1) & (2.68) & (0.012) \\
Constant & 0.73*** & 835.6*** & 679.3*** & 82.8*** & -0.017 \\
      & (0.063) & (93.8) & (80.0) & (6.14) & (0.017) \\
      &       &       &       &       &  \\
\midrule
Observations & 5062  & 5085  & 5085  & 2490  & 2653 \\
R-sq  & 0.021 & 0.011 & 0.010 & 0.041 & 0.232 \\
Control Mean & 0.73  & 829.6 & 670.9 & 92.7  & 0.092 \\
\midrule
DOW FE & YES   & YES   & YES   & YES   & YES \\
Branch FE & YES   & YES   & YES   & YES   & YES \\
Number of pawns FE & YES   & YES   & YES   & YES   & YES \\
Number of arms FE & YES   & YES   & YES   & YES   & YES \\
\bottomrule
\bottomrule
\end{tabular}%