### 2020-05-30 presentations/beamer_pola.tex

\documentclass[9pt]{beamer}

\input{presentations/preamble_beamer.tex}

%\setbeamercolor{section in head/foot}{bg=darkcrimsonred}
\setbeamersize{text margin left=11pt, text margin right=11pt}
%\setbeamertemplate{section in toc}[square]

\begin{document}

\title{Frequent Payments as a Commitment Device?}

%\subtitle{VERY PRELIMINARY!}

\author[]{
Isaac Meza\inst{1}\and Joyce Sadka\inst{2}  \and Enrique Seira\inst{2}}

\institute{ \inst{1} University of California, Berkeley \and
  \inst{2}  Instituto Tecnológico Autónomo de México}

\date{Oct, 2019\\\;\;\\\;\;}

\begin{frame}[c, noframenumbering]%{\phantom{title page}}
% The \phantom{title page} is a kludge to get the red bar on top
\titlepage
\end{frame}

\begin{frame}{Motivation}
\begin{itemize}
\item How the collaboration started
    \begin{itemize}
        \item We were contacted by the organization about their contract.
        \item Conjectured that frequent payments would lead to higher recovery.

    \end{itemize}
    \pause
\vfill \item The standard contract:
    \begin{itemize}
        \item Accepted gold jewelry as collateral, gave you 70\% of the value of gold in cash.
        \item 7\% Monthly interest in outstanding debt.
        \item \textbf{3 month term}. You could pay principal+interest anytime with no penalty and recover your item. 
        \item if in 90 (+15 days grace) you did not do this, you lose your item.
        \item A bit more than 1/3 lost their pawn.
        \item Only 33\% paid before day 85th in our data/
    \end{itemize}
    \pause
    \end{itemize}

\vfill
\end{frame}

\begin{frame}{2 Questions}
        \begin{enumerate}
 \item Does forcing clients to \textbf{pay monthly} increase payment (decrease default)?
    \begin{itemize}
        \item \textbf{Higher frequency has costs}: $\uparrow$ transaction costs, $\downarrow$  welfare (with neo-classical consumers).
        \item But may help present biased consumers. 
    \end{itemize}

    \vfill
    \pause
    \item \textbf{Is there demand for commitment?}
        \begin{itemize}
            \item Would clients demand it if given a choice between monthly payments vs at anytime in 3 months?
            \item Who demands it?
            \item Neoclassical consumers would \textbf{not} tie their hands for free.
        \end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}{What ``punishment'' to use for late payment?}

\begin{itemize}
    \item There needs to be a punishment from not paying in order to have commitment.
    \vfill
    \item But punishment if it is too hard, and there are random shocks to income, even present biased consumers will not demand it.
    \vfill 
    \item Tried with 2 types of punishment:
    \begin{itemize}
        \item Monetary fee: $4\%$ of missed payment.
        \item Psychological cost of lying: made promise very salient.
    \end{itemize}
\end{itemize}
\end{frame}

% LOS SIGUIENTE ES PARA QUE PONGA SECCIONES Y DIVISIONES ENTRE SECCIONES
	\setbeamertemplate{section in toc}{\hypersetup{linkcolor=.}\colorbox{lightash}{\textcolor{white}{\inserttocsectionnumber}}\enskip\inserttocsection}

	\begin{frame}[t]{Steps}
	\hfill
	\parbox[t]{.95\textwidth}{
		\begin{minipage}[c][0.75\textheight]{\textwidth}
			\tableofcontents[hideallsubsections]	
	\end{minipage}}
	\hfill
\end{frame}

\AtBeginSection[]
{
\begin{frame}[t]{Steps}
\hfill
\parbox[t]{.95\textwidth}{
	\begin{minipage}[c][0.75\textheight]{\textwidth}
		\tableofcontents[currentsection,hideallsubsections,sectionstyle=show/shaded]  				
\end{minipage}}
\hfill
\end{frame}
}

%EMPIEZAN LAS SECCIONES

\section{Experiment description}

\begin{frame}{Experimental Arms}
     5 Arms:
        \begin{itemize}
            \item \textbf{Control} = status quo pay any time in 90 days.
            \vfill
            \item \textbf{Frequent payment treatments}: 4 arms in a 2x2 design: 
            \vfill
            \begin{itemize}
                \item ``Forced'' to have the monthly payment contract vs given choice.
                \vfill
                \item  Break promise vs penalty fee (4\%) if delinquent on payment.
            \end{itemize}
            \vfill
            \item Randomization at branch-day level.
                \begin{itemize}
                    \item No difference in number of contracts by arm.
                \end{itemize}
        \end{itemize}
    \vfill
    \begin{center}
        \centering
        \includegraphics[width=0.70\textwidth]{Figuras/exp_arms.PNG}
    \end{center}
\end{frame}

\begin{frame}{Explanation in ``choice'' days}
    \begin{center}
        \includegraphics[width=\textwidth]{Figuras/micas.pdf}
    \end{center}
\end{frame}

\begin{frame}{Branches and Timeline}
    \begin{itemize}
        \item We worked on 6 branches during 2012
        \item We observe transactions for about 30 days before and 6 months after the experiment ends.
    \end{itemize}
    \begin{center}
        \includegraphics[width=0.80\textwidth]{Figuras/timeline_suc_exp_extended.pdf}
    \end{center}
\end{frame}

\begin{frame}{Data}

\begin{itemize}
    \item Admin data:
    \begin{itemize}
        \item Date of pawning
        \item Value of item --> loan size
        \item Dates of payments and amounts, 
        \item Whether borrower recovered item
        \item Borrower id
    \end{itemize}
    \vfill
    \item Survey data: Own 2 page questionnaire at the branch before pawning
    \begin{itemize}
        \item Some demographics
        \item Vulnerability
        \item Cost of going to branch
        \item 3 ``self control'' questions
        \item ... others.
    \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{Survey questions}
    \begin{enumerate}
        \item Family
            \begin{itemize}
                \item Common asks - Is it common family or friends ask you for money?
            \end{itemize}
        \item Income 
        \begin{itemize}
            \item In the last six months, did you need money to pay...?
        \end{itemize}
        \item Self-Control
        \begin{itemize}
            \item Makes budget - Do you budget your monthly expenses?
            \item Present bias - Prefers 100 tomorrow over 250 in a month AND 150 in 4 months over 100 in 3 months
            \item Reminder - Would you like to receive free reminders on your cell phone?
            \item Tempt - Has it happened to you that you spend more than you would like because you fall into temptation?
        \end{itemize}
        \item Experience
        \begin{itemize}
            \item Rosca - Do you participate in a \textit{rosca} or in a savings account?
            \item Pawn before - Have you pawn before?
        \end{itemize}
        \item Other
            \begin{itemize}
            \item Low time - Below median in reported time to arrive to the pawn shop.
            \item Low cost - Below median in reported cost spent to arrive to the pawn shop
            \item Prob recovery - From 0 to 100 mark with an X how sure you are to recover your garment
            \item Stressed - How often do you feel stressed by your household economic situation? - 0 when answered `never' and 1 otherwise
        \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}{Summary Stats}
    \begin{table}
        \begin{center}
        \tiny{\input{./Tables/SS.tex}}
        \end{center}
    \end{table}
\end{frame}

\begin{frame}{SS Histograms}

\begin{figure}[H]
    \label{nochoicehte_takeup_pred}
    \begin{center}
     \begin{subfigure}{0.40\textwidth}
        \caption{In the last six months, did you need money to pay...?}
        \centering
        \includegraphics[width=\textwidth]{Figuras/perc_utilities.pdf}
    \end{subfigure}
     \begin{subfigure}{0.5\textwidth}
        \caption{Cost and Time to arrive to pawn shop}
        \centering
        \includegraphics[width=\textwidth]{Figuras/cost_time_scatter_hist.png}
    \end{subfigure}
    \end{center}
    \end{figure}

\end{frame}

\section{Results on Payments and Recovery}

\begin{frame}{Recovering Piece}

    \begin{center}
        \includegraphics[width=.70\textwidth]{Figuras/te_graph_des_c.pdf}
    \end{center}
\end{frame}

\begin{frame}{Percentage of Loan paid}
    \begin{center}
        \includegraphics[width=.70\textwidth]{Figuras/te_graph_sum_porcp_c.pdf}
    \end{center}
\end{frame}

\begin{frame}{Days to recover piece | recovering}
    \begin{center}
        \includegraphics[width=.70\textwidth]{Figuras/te_graph_dias_al_desempenyo.pdf}
    \end{center}
\end{frame}

\begin{frame}{Reincidence}
    \begin{center}
        \includegraphics[width=.70\textwidth]{Figuras/te_graph_reincidence.pdf}
    \end{center}
\end{frame}

\begin{frame}[label = hte]{TE heterogeneity in Recovering}
    \begin{center}
        \includegraphics[width=.80\textwidth]{Figuras/he_dist_des_c.pdf}
    \end{center}

 Based on \hyperlink{GRF}{\beamerbutton{GRF}}   
\end{frame}

\begin{frame}{A proxy for financial cost to client}

Financial cost is defined as: sum of payments + cost of piece if not recovered

    \begin{center}
        \includegraphics[width=.7\textwidth]{Figuras/hist_fc.pdf}
    \end{center}
\end{frame}

\begin{frame}{Effect of Frequency on Financial Cost}

\begin{itemize}
    \item Financial cost decreases by 350 pesos, from a mean of 3,200.
\end{itemize}
\vspace{.2in}
\vfill
    \begin{center}
        \includegraphics[width=.9\textwidth]{Figuras/FC_table.png}
    \end{center}
\end{frame}

\section{Results on Demand for Commitment}

\begin{frame}{Fit: Machine learning models}
\begin{itemize}
    \item 87\% accuracy: we can predict take up well.
    \vfill
\end{itemize}
    \begin{center}
    \input{./Tables/oos_pago_frec_vol_fee.tex}
    \end{center}
\end{frame}

\begin{frame}{10\% take up with fee: who are they?}
    \begin{center}
        \includegraphics[width=.90\textwidth]{Figuras/pago_frec_vol_fee_interactions_rf.pdf}
    \end{center}
\end{frame}

\begin{frame}{``Benefit'' positively correlated with predicted take-up}

\begin{itemize}
    \item Estimate take-up \alert{in Choice fee group} as a function of X's: $f(\beta,X)$.
    \item Predict take up using $f(\beta,X)$ in \alert{No-choice group}.
    \item relate predicted take up to effect they really had in \alert{No-choice group}.
\end{itemize}

\begin{figure}[H]
    \begin{center}
     \begin{subfigure}{0.45\textwidth}
        \caption{Recovery vs predicted take-up}
        \centering
        \includegraphics[width=\textwidth]{Figuras/takeup_he_pro_2_pago_frec_vol_fee_des_c.pdf}
    \end{subfigure}
     \begin{subfigure}{0.45\textwidth}
        \caption{FC vs predicted take-up}
        \centering
        \includegraphics[width=\textwidth]{Figuras/takeup_he_pro_2_pago_frec_vol_fee_fc_survey.pdf}
    \end{subfigure}
    \end{center}
    \end{figure}

\end{frame}

\section{Next steps}

\begin{frame}{Potential projects and remaining questions}

\begin{enumerate}

    \item Current project:
    \begin{enumerate}
        \item Get historical data to see results by different types of people.
        \item Replicate experiment in 2 or 3 branches, with improved design.
        \begin{itemize}
            \item Ask preference to all, then randomize.
            \item Better satisfaction survey.
            \item Measure preferences better at baseline, and expectations.
            \item Cross check with bureau and savings.
            \item Measure mistakes in take up
        \end{itemize}
    \end{enumerate}

    \vfill
    \item Impact of different collateral requirements:
    \begin{itemize}
        \item Pilot 65\% and 75\% collateral and measure take up, recovery and profits
    \end{itemize}

    \vfill
    \item Reporting to Credit Bureau (pilot):
    \begin{itemize}
        \item As a way to impose discipline (ex-ante vs ex-post)
        \item It can be voluntary
    \end{itemize}

    \vfill
    \item Optimal auctions to sell the goods in a webpage.

    \vfill
    \item Tracking by cellphone to screen for bank.

\end{enumerate}
\end{frame}

\section{Appendix}

\begin{frame}{When do borrowers pay?}
    \begin{center}
        \includegraphics[width=\textwidth]{Figuras/hist_payments.pdf}
    \end{center}
\end{frame}

\begin{frame}{Histograms Percentage Paid}
    \begin{center}
        \includegraphics[width=.9\textwidth]{Figuras/hist_perc_payment.pdf}
    \end{center}
\end{frame}

\begin{frame}{Number of Payments}
    \begin{center}
        \includegraphics[width=.70\textwidth]{Figuras/te_graph_num_p.pdf}
    \end{center}
\end{frame}

\begin{frame}{Some determinants of TE - The nature of heterogeneity}
    \begin{center}
        \includegraphics[width=.9\textwidth]{Figuras/he_int_vertical_des_c_pro_2.pdf}
    \end{center}
\end{frame}

\begin{frame}{Characteristics of present bias}
\begin{table}[H]
\label{pb_chars}
\begin{center}
\tiny{\input{./Tables/pb_chars.tex}}
\end{center}
\end{table}
\end{frame}

\begin{frame}{30\% take up with promise: Different relation with self control}
    \begin{center}
        \includegraphics[width=.90\textwidth]{Figuras/pago_frec_vol_promise_interactions_rf.pdf}
    \end{center}
\end{frame}

\begin{frame}{Exit survey}
    \begin{center}
    \tiny
    \input{./Tables/SS_exit.tex}
    \end{center}
\end{frame}

\begin{frame}[label = GRF]{Generalized Random Forests}

A method for non-parametric statistical estimation based on RF
for estimating $\theta(x)$ that can be identified as the solution to a set of local moment equations:

\[\mathbb{E}[\psi(Z,\theta,h(x,W))\;|\;X=x]=0\]

where $\psi$ is a known score function and $h$ is an unkown nuisance function also estimated from data.\\

\vspace{5mm}

The main application involves HTE. To achieve this, the method seeks trees that, when combined, produces leaves that maximizes heterogeneity. In other words, it splits the data that produce the biggest difference in TE across leaves, by still producing accurate estimates of the TE. Moreover, the splits are made honest by taking the training data and splitting it into two subsamples: a splitting subsample and an estimating subsample. The estimates being calculated on the latter. \hyperlink{hte}{\beamerbutton{Back}}\\

\end{frame}

\begin{frame}{Honest Causal Tree construction}
Given a dataset with an outcome $Y$, covariates $X$, and a randomized condition $W$ that takes on the value of 0 for control and 1 for treatment:

\begin{enumerate}
\item Split the data into subsample $I$ and subsample $J$

\item Train a decision tree on subsample $I$ predicting $Y$ from $X$, with the requirement that each terminal node has at least $k$ observations from each condition in subsample $J$

\item Apply the decision tree constructed on subsample $I$ to subsample $J$

\item At each terminal node, get the mean of predictions for the $W = 1$ cases from subsample $J$ and subtract the mean of predictions for the $W = 0$ cases from subsample $J$; the resulting difference is the estimated treatment effect.

\end{enumerate}

This is then applied to a causal random forest by using bagging or bootstrapping

\end{frame}
\end{document}