### 2020-06-18 Tables/transcribed.tex

% Table generated by Excel2LaTeX from sheet 'transcribed'
\begin{tabular}{cl}
\toprule
      & \textbf{Baseline Survey} \\
\midrule
\midrule
1     & \textbf{Your pawn was:} \\
      & (a) Inherite, (b) a gift, (c) bought by me, (d) lend to me, (e) other \_\_\_\_\_\_\_\_\_\_\_\_ \\
2     & \textbf{Mark with an "X" in the line below how likely is that you recover your pawn. } \\
      & \textbf{Where 0 is impossible and 100 is completely certain} \\
      &  \\
      &  \\
3     & \textbf{How much would you sell the item you want to pawn for?       \_\_\_\_\_\_\_\_\_\_\_\_\_\_ pesos} \\
4     & \textbf{Gender      } \\
5     & \textbf{Age} \\
6     & \textbf{Civil Status } \\
      & (a) married, (b) single, (c) divorced, (d) widowed \\
7     & \textbf{Work status} \\
      & (a) employed, (b) own business, (c) houseshores, (d) don't work, (e) retired, (f) study \\
8     & \textbf{Education} \\
      & (a) no formal education, (b) primary, (c) middle school, (d) highschool, (e) more than highschool \\
9     & \textbf{In the last month, did a friend or family member asked you for money?} \\
      & (a) yes  (b) no \\
10    & \textbf{What would you like to have: 100 pesos tomorrow or 150 pesos in one month?} \\
11    & \textbf{How often do you feel stressed by your economic situation?} \\
      & (a) always, (b) very often, (c) sometimes, (d) never \\
12    & \textbf{What is the main reason you want to pawn?} \\
      & (a) Need the money because somebody in my family lost his/her job \\
      & (b) Need the money to pay for a sickness in the family \\
      & (c) Need the money for an urgent expense \\
      & (d) Need the money for some non urgent expense. \\
13    & \textbf{How stressed do you feel from the situation that led to to pawn?} \\
      & (a) very stressed, (b) somwhat stressed, (c) a little stressed, (d) not stressed  \\
14    & \textbf{In 3 months, I expect to have a  \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_ situation} \\
      & (a) better, (b) similar,  (c) worse \\
15    & \textbf{Have you panwned before?} \\
      & (a) yes  (b) no \\
16    & \textbf{How many times have you pawned on a Lender P branch?} \\
      & (a) NO\_\_\_    (b)  1-2 times \_\_\_    (c) 3-5 times\_\_\_\_   (d) More than 5\_\_\_\_ \\
17    & \textbf{If you are saving money and a family member wants to use it for something } \\
      & (a) I would only give him the money for an urgent expenze \\
      & (b) I would give him the money even if it was not an urgent expense \\
      & (c) I would not give him/her the money regardless \\
      & (d) No one would ask me for my money \\
18    & \textbf{Do you make an expenses budget for the month ahead of time?} \\
      & (a) always, (b) very often, (c) sometimes, (d) never \\
19    & \textbf{Do you have other items you could pawn?} \\
      & (a) yes  (b) no \\
20    & \textbf{Do you have savings?} \\
      & (a) yes  (b) no \\
21    & \textbf{Do you participate in a ROSCA?} \\
      & (a) yes  (b) no \\
22    & \textbf{Is it common that family or friends ask for money?} \\
      & (a) yes  (b) no \\
23    & \textbf{How much did you spend to come to the branch today?    \$\_\_\_\_\_\_\_\_\_\_\_\_\_\_ pesos} \\
24    & \textbf{How much time does it usually take to come to this branch?    \_\_\_\_\_\_\_\_\_\_\_} \\
25    & \textbf{How much does your family spend in a normal week?   \$\_\_\_\_\_\_\_\_\_\_\_\_\_\_ pesos} \\
26    & \textbf{How much do you manage to save in a normal week?   \$\_\_\_\_\_\_\_\_\_\_\_\_\_\_ pesos} \\
27    & \textbf{Does it happen to you that you spend more than you wanted because you fall into temptation?} \\
      & (a) never, (b) almost never, (c) sometimes, (d) very often \\
28    & \textbf{In the last 6 months, has it happened that at some point you lacked money to pay} \\
      & (a) rent?    (b) food    (c)food   (d) medicine  (e) electricity   (f) heating   (g) telephone    (i) water \\
29    & \textbf{What would you like to have: 100 pesos in 3 months or 150 pesos in four months?} \\
30    & \textbf{Would you like to receive (free) reminders for upcomming payments?} \\
      & (a) yes  (b) no \\
\bottomrule
\end{tabular}%