### 2021-11-13 choice_benefit_summary.tex
Learning by not doing regressions

 \documentclass[oneside,11pt]{article}

\input{preamble.tex}
%%% HELPER CODE FOR DEALING WITH EXTERNAL REFERENCES
\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{
    \externaldocument{#1}
    \addFileDependency{#1.tex}
    \addFileDependency{#1.aux}
}

%\myexternaldocument{OA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DOCUMENT
\begin{document}

We estimate demand for commitment with a Gradient Boosted Classification algorithm (accuracy 97\% in-sample and 90\% out-of-sample).

Now we want to identify who are the ones who benefit more from being forced, and those who would demand commitment.

Based of the propensity score, we predict the two categories : choosers \& non-choosers, such that the proportions match the observed data.

\begin{figure}[H]
    \caption{}
    \label{}
    \begin{center}
    %\begin{subfigure}{0.4\textwidth}
    %    \caption{Financial cost}
    %    \centering
    %    \includegraphics[width=\textwidth]{Figuras/he_dist_fc_admin_disc_pro_2.pdf}
    %\end{subfigure}
    \begin{subfigure}{0.475\textwidth}
        \caption{Effective cost/loan benefit}
        \centering
        \includegraphics[width=\textwidth]{Figuras/cdf_predchoose_tau_eff.pdf}
    \end{subfigure}
    \begin{subfigure}{0.475\textwidth}
        \caption{Repayment}
        \centering
        \includegraphics[width=\textwidth]{Figuras/cdf_predchoose_tau_des.pdf}
    \end{subfigure}

    \end{center}
     \scriptsize    CDF of the heterogeneous treatment effect for choosers vs non-choosers. The dotted line below indicates the points where the difference in the distributions is significant.  Instead of a single global null hypothesis (that the two CDFs are identical), there is a continuum of individual null hypotheses of CDF equality at each point. The methodology was proposed by \cite{GOLDMAN2018143}.

          %\footnotesize{ \textit{Do file: }  \texttt{benefit\_choice.do}}
\end{figure}

We find a surprising \emh{significant} stochastic dominance between the two CDF's.

\newpage

\section{Observational Study}

\subsection*{Paternalism  with learning : We generate more learning through forcing than through choice.}

It is possible that the wrong people choose and they learn not through choosing, but rather through not choosing (and it did not work). We then partition our sample in the observational study, based on whether you committed the first time, and we want to show that learning happened in the ``wrong group" : not choosers.\\

We want to show that the TE is localized within one of the groups in that endogenous split. With that in mind, we estimate the following two-stage specification.

\begin{equation}
    \mathbbm{1}(\text{Had FP})_{ijt} = \alpha_i + \gamma_t + \beta \mathbbm{1}(\text{FP Avail})_{ijt}  + \epsilon_{ijt}
    \label{eqn:fs}
\end{equation}

\begin{equation}
    \mathbbm{1}(\text{Has FP})_{ijt} = \widehat{\alpha_i }+ \widehat{\gamma_t} + \widehat{\beta }(\widehat{\text{Had FP }})_{ijt}  + \epsilon_{ijt}
    \label{eqn:2s}
\end{equation}

where $i,j,t$ index client, branch, and week respectively. $\mathbbm{1}(\text{Has FP})_{ijt}$ is an indicator for client $i$ pawning in branch $j$ in week $t$ using a FP contract, given that both FP and traditional contracts were available at the branch at the time of pawning, so that there is a choice to be made among these. We instrument $\mathbbm{1}(\text{Had FP previously})_{ijt}$ with the \textit{availability} of the contract in branch $j$  when the client previously went to pawn the immediately prior pawn with respect to $t$, $\mathbbm{1}(\text{FP Avail})_{ijt}$. However, the second stage only uses the subsample of the not choosers.

\begin{table}[H]
\caption{Outcome in 2S : Choose FP contract}
\label{iv_pf}
\begin{center}
\scriptsize{\input{./Tables/iv_reg_pago_fijo_detailed.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{iv\_reg\_detailed.do}
\end{table}

Thus the learning mechanism is due to the not choosers having failed.

\newpage

\subsection*{The results are robust when dropping `` murky" cases:}

\begin{table}[H]
\caption{Keeping only first two observations}
\begin{center}
\scriptsize{\input{./Tables/iv_reg_pago_fijo_detailed_twoca.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{iv\_reg\_detailed.do}
\end{table}

\begin{table}[H]
\caption{Collapsing observations}
\begin{center}
\scriptsize{\input{./Tables/iv_reg_pago_fijo_detailed_colla.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{iv\_reg\_detailed.do}
\end{table}

\subsection*{Learning is stronger after some experience.}

\begin{table}[H]
\caption{Interacting with number of decision epochs}
\begin{center}
\scriptsize{\input{./Tables/iv_reg_pago_fijo_detailed_inter.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{iv\_reg\_detailed.do}
\end{table}

\subsection*{What about the impact of the previous outcome (default) on learning?}

We estimate the following specification, instrumenting the previous contract with the availability of the product as before.

\begin{align*}
    \mathbbm{1}(\text{Has FP})_{ijt} &= \alpha_i + \gamma_t + \beta_1\mathbbm{1}(\text{Had FP})_{ijt} \\
    & \qquad \quad + \beta_2\mathbbm{1}(\text{Defaulted})_{ijt} \\
    & \qquad \quad  + \beta_3\mathbbm{1}(\text{Had FP})_{ijt}\mathbbm{1}(\text{Defaulted})_{ijt}   + \epsilon_{ijt}
\end{align*}

\begin{table}[H]
\caption{Interaction with default in the past}
\begin{center}
\scriptsize{\input{./Tables/iv_reg_pago_fijo_def_detailed.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{iv\_reg\_intdef\_detailed.do}
\end{table}

And as before we can also get rid of the `murky' cases:

% \begin{table}[H]
% \caption{Keeping only first two observations}
% \begin{center}
% \scriptsize{\input{./Tables/iv_reg_pago_fijo_def_twocases.tex}}
% \end{center}
%  \scriptsize

% %\textit{Do file: } \texttt{iv\_reg\_intdef\_detailed.do}
% \end{table}

\begin{table}[H]
\caption{Collapsing}
\begin{center}
\scriptsize{\input{./Tables/iv_reg_pago_fijo_def_collapsed.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{iv\_reg\_intdef\_detailed.do}
\end{table}

\section{Learning with experimantal data}

\begin{table}[H]
\caption{Summary statistics table of learning by not doing (OLS)}
\begin{center}
\scriptsize{\input{./Tables/SS_learning_OLS.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{learning\_exp.do}
\end{table}

This might look discouraging, but what if we exploit individual variation?

The above means can be recovered from the following specification

\begin{align*}
    \text{Choose Fee}_{i} &=  \alpha +  \beta_1\mathds{1}(\text{Default past})_{i} +\sum_{j=1}^{4}\beta_{2,j}\mathds{1}(\text{Past contract}=j)_{i} \\
   &\qquad\quad  +\sum_{j=1}^{4}\beta_{3,j}\mathds{1}(\text{Past contract}=j)_{i}\mathds{1}(\text{Default past})_{i} + \epsilon_i
\end{align*}
for those individuals $i$ that had the option to choose after experiencing contract $j$, this could be control, forced fee, or choice.

So what if we add fixed effects to the above? In order to capture variation within individuals in line with our previous specifications in the observational study.

\subsection*{We find there is learning, and also learning by not doing!}

\begin{table}[H]
\caption{Summary statistics table of learning by not doing (FE)}
\begin{center}
\scriptsize{\input{./Tables/SS_learning_FE.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{learning\_exp.do}
\end{table}

\begin{table}[H]
\caption{Learning by not doing (experimental)}
\begin{center}
\scriptsize{\input{./Tables/learning_exp.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{learning\_exp.do}
\end{table}

\newpage

\section{LATE}

\begin{table}[H]
\caption{}
\label{tot_tut}
\begin{center}
\scriptsize{\input{./Tables/tot_tut_iv.tex}}
\end{center}
 \scriptsize
%\textit{Do file: } \texttt{tot\_tut.do}
\end{table}

\begin{figure}[H]
    \caption{Bootstrap inference for the difference between TOT-TUT}
    \label{}
    \begin{center}
    \begin{subfigure}{0.45\textwidth}
        \caption{Simple}
        \centering
        \includegraphics[width=\textwidth]{Figuras/tot_tut_btsp1.pdf}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
        \caption{Admin controls}
        \centering
        \includegraphics[width=\textwidth]{Figuras/tot_tut_btsp2.pdf}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
        \caption{Survey controls}
        \centering
        \includegraphics[width=\textwidth]{Figuras/tot_tut_btsp3.pdf}
    \end{subfigure}

    \end{center}
     \scriptsize   
          %\footnotesize{ \textit{Do file: }  \texttt{tot\_tut.do}}
\end{figure}

\begin{figure}[H]
     \caption{TOT-TUT Histogram}
    \label{}
    \begin{center}
    \begin{subfigure}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{Figuras/dif_tot_tut.pdf}
    \end{subfigure}
    \end{center}
    \scriptsize
        Histogram for the difference of the conditional (heterogeneous) TOT-TUT
          %\footnotesize{ \textit{Do file: }  \texttt{tot\_tut\_insforest.do}}
\end{figure}

\begin{figure}[H]
     \caption{TOT-TUT CDF}
    \begin{center}
    \begin{subfigure}{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{Figuras/cdf_tot_tut.pdf}
    \end{subfigure}
    \end{center}
    \scriptsize
        ECDF of the conditional (heterogeneous) TOT \& TUT, alogn with the difference and a test for the difference in means (which is significant).
          %\footnotesize{ \textit{Do file: }  \texttt{tot\_tut\_insforest.do}}
\end{figure}

We test the exclusion restriction

\begin{figure}[H]
     \caption{Exclusion restriction}
    \begin{center}
    \begin{subfigure}{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{Figuras/exclusion_restriction.pdf}
    \end{subfigure}
    \end{center}
    \scriptsize

          %\footnotesize{ \textit{Do file: }  \texttt{exclusion\_restriction.do}}
\end{figure}

\newpage
\clearpage
\bibliographystyle{authordate1}
%\bibliographystyle{amsalpha}
%\bibliographystyle{AER}

\bibliography{References_aux.bib}

\end{document}