### 2023-02-28 Tables/decomposition_main_te_pres1.tex

% Table generated by Excel2LaTeX from sheet 'decomposition_main_te_pres1'
\begin{tabular}{lcccccc}
\toprule
      &       & \multicolumn{5}{c}{Components of FC} \\
\cmidrule{3-7}      & FC    & \textcolor[rgb]{ .749,  .749,  .749}{Int. pymnt} & \textcolor[rgb]{ .749,  .749,  .749}{Fee pymnt} & \textcolor[rgb]{ .749,  .749,  .749}{Princ. pymnt} & \textcolor[rgb]{ .749,  .749,  .749}{Lost pawn val} & Default \\
\midrule
      & (1)   & \textcolor[rgb]{ .749,  .749,  .749}{(2)} & \textcolor[rgb]{ .749,  .749,  .749}{(3)} & \textcolor[rgb]{ .749,  .749,  .749}{(4)} & \textcolor[rgb]{ .749,  .749,  .749}{(5)} & (6) \\
\midrule
\midrule
Forced cmit & \boxit{0.5in} -379.7*** & \textcolor[rgb]{ .749,  .749,  .749}{-157.3***} & \textcolor[rgb]{ .749,  .749,  .749}{32.1***} & \textcolor[rgb]{ .749,  .749,  .749}{-0.57} & \textcolor[rgb]{ .749,  .749,  .749}{-254.5**} & \boxit{0.5in} -0.065*** \\
      & (111.4) & \textcolor[rgb]{ .749,  .749,  .749}{(34.9)} & \textcolor[rgb]{ .749,  .749,  .749}{(1.43)} & \textcolor[rgb]{ .749,  .749,  .749}{(3.03)} & \textcolor[rgb]{ .749,  .749,  .749}{(104.8)} & (0.023) \\
Choice cmit & -84.9 & \textcolor[rgb]{ .749,  .749,  .749}{-24.9} & \textcolor[rgb]{ .749,  .749,  .749}{1.34**} & \textcolor[rgb]{ .749,  .749,  .749}{-3.98} & \textcolor[rgb]{ .749,  .749,  .749}{-61.4} & -0.025 \\
      & (114.6) & \textcolor[rgb]{ .749,  .749,  .749}{(38.4)} & \textcolor[rgb]{ .749,  .749,  .749}{(0.54)} & \textcolor[rgb]{ .749,  .749,  .749}{(2.47)} & \textcolor[rgb]{ .749,  .749,  .749}{(109.2)} & (0.021) \\
      &       & \textcolor[rgb]{ .749,  .749,  .749}{} & \textcolor[rgb]{ .749,  .749,  .749}{} & \textcolor[rgb]{ .749,  .749,  .749}{} & \textcolor[rgb]{ .749,  .749,  .749}{} &  \\
\midrule
Observations & 6304  & \textcolor[rgb]{ .749,  .749,  .749}{6304} & \textcolor[rgb]{ .749,  .749,  .749}{6304} & \textcolor[rgb]{ .749,  .749,  .749}{6304} & \textcolor[rgb]{ .749,  .749,  .749}{6304} & 6304 \\
R-squared & 0.007 & \textcolor[rgb]{ .749,  .749,  .749}{0.022} & \textcolor[rgb]{ .749,  .749,  .749}{0.151} & \textcolor[rgb]{ .749,  .749,  .749}{0.003} & \textcolor[rgb]{ .749,  .749,  .749}{0.007} & 0.013 \\
Control Mean & 1851.0 & \textcolor[rgb]{ .749,  .749,  .749}{545.9} & \textcolor[rgb]{ .749,  .749,  .749}{0} & \textcolor[rgb]{ .749,  .749,  .749}{5.82} & \textcolor[rgb]{ .749,  .749,  .749}{1305.1} & 0.44 \\
\bottomrule
\bottomrule
\end{tabular}%