### 2023-10-16 Tables/Attrition.tex
cleaning of project folder - build replication files

% Table generated by Excel2LaTeX from sheet 'Attrition'
\begin{tabular}{lcccc}
\toprule
      &       & \multicolumn{3}{c}{Commitment arms} \\
\cmidrule{3-5}      & Control & Forced & Choice & p-value \\
\midrule
\midrule
Number of branch-day pawns & 35    & 32    & 37    & 0.46 \\
      & (3.5) & (2.7) & (3.1) &  \\
\multicolumn{1}{p{12.82em}}{Ended up pawning} & 0.98  & 0.97  & 0.97  & 0.62 \\
      & (0.01) & (0.01) & (0.01) &  \\
Survey response rate & 0.79  & 0.76  & 0.77  & 0.62 \\
      & (0.02) & (0.02) & (0.02) &  \\
\midrule
Obs   & 1770  & 1954  & 2580  &  \\
\bottomrule
\bottomrule
\end{tabular}%