### 2023-10-16 Tables/old/exp_arms.tex

% Table generated by Excel2LaTeX from sheet 'exp_arms'
\begin{tabular}{p{7.68em}ccccc}
\toprule
\multicolumn{1}{r}{} &       & \multicolumn{2}{c}{\cellcolor[rgb]{ .675,  .725,  .792}Fee arms} & \multicolumn{2}{c}{\cellcolor[rgb]{ .851,  .882,  .949}Promise arms} \\
\midrule
\multicolumn{1}{r}{} & \multicolumn{1}{p{4.045em}}{Control} & \multicolumn{1}{p{3.365em}}{Forced-fee} & \multicolumn{1}{p{4.045em}}{Choice} & \multicolumn{1}{p{4.045em}}{Forced} & \multicolumn{1}{p{4.045em}}{Choice } \\
\cmidrule{2-6}\rowcolor[rgb]{ .706,  .776,  .906} Randomization arms & \cellcolor[rgb]{ .851,  .882,  .949}2585 & \cellcolor[rgb]{ .851,  .882,  .949}2465 & \cellcolor[rgb]{ .851,  .882,  .949}3406 & \cellcolor[rgb]{ .851,  .882,  .949}2143 & \cellcolor[rgb]{ .851,  .882,  .949}2755 \\
Surveys & 2036  & 1907  & 2710  & 1757  & 2216 \\
\rowcolor[rgb]{ .706,  .776,  .906} Ended up pawning & \cellcolor[rgb]{ .851,  .882,  .949}1984 & \cellcolor[rgb]{ .851,  .882,  .949}1840 & \cellcolor[rgb]{ .851,  .882,  .949}2634 & \cellcolor[rgb]{ .851,  .882,  .949}1724 & \cellcolor[rgb]{ .851,  .882,  .949}2186 \\
\bottomrule
\bottomrule
\end{tabular}%