### 2023-10-21 Tables/Attrition.tex
stabilize cleaning files & internal validity analysis for number of pawns balance

% Table generated by Excel2LaTeX from sheet 'Attrition'
\begin{tabular}{lcccc}
\toprule
      &       & \multicolumn{3}{c}{Commitment arms} \\
\cmidrule{3-5}      & Control & Forced & Choice & p-value \\
\midrule
\midrule
Number of branch-day pawns & 31    & 31    & 36    & 0.19 \\
      & (2)   & (2.1) & (2)   &  \\
\multicolumn{1}{r}{median} & 27    & 28    & 31    & 0.15 \\
\midrule
Number of branch-day borrowers & 21    & 23    & 25    & 0.1 \\
      & (1.2) & (1.4) & (1.3) &  \\
\multicolumn{1}{r}{median} & 19    & 21    & 21    & 0.46 \\
\midrule
\multicolumn{1}{p{15.82em}}{Ended up pawning} & 0.98  & 0.97  & 0.97  & 0.62 \\
      & (0.01) & (0.01) & (0.01) &  \\
Survey response rate & 0.79  & 0.76  & 0.77  & 0.62 \\
      & (0.02) & (0.02) & (0.02) &  \\
\midrule
Obs   & 101   & 97    & 129   &  \\
\bottomrule
\bottomrule
\end{tabular}%