### 2023-11-15 Tables/Attrition.tex
limited and balanced attrition - analysis for internal validity

% Table generated by Excel2LaTeX from sheet 'Attrition'
\begin{tabular}{lcccc}
\toprule
      &       & \multicolumn{3}{c}{Commitment arms} \\
\cmidrule{3-5}      & Control & Forced & Choice & p-value \\
\midrule
\midrule
Number of branch-day pawns & 31    & 31    & 36    & 0.21 \\
      & (5.5) & (5.4) & (6.4) &  \\
\multicolumn{1}{r}{median} & 27    & 28    & 31    & 0.15 \\
\midrule
Number of branch-day borrowers & 21    & 23    & 25    & 0.24 \\
      & (3.3) & (3.8) & (4.3) &  \\
\multicolumn{1}{r}{median} & 19    & 21    & 21    & 0.46 \\
\midrule
Obs   & 101   & 97    & 129   &  \\
\midrule
\midrule
\multicolumn{1}{p{15.82em}}{Ended up pawning} & 0.98  & 0.97  & 0.97  & 0.62 \\
      & (0.01) & (0.01) & (0.01) &  \\
Survey response rate & 0.79  & 0.76  & 0.77  & 0.62 \\
      & (0.02) & (0.02) & (0.02) &  \\
\midrule
Obs   & 1770  & 1954  & 2580  &  \\
\bottomrule
\bottomrule
\end{tabular}%