### 2024-05-24 Tables/Attrition.tex

% Table generated by Excel2LaTeX from sheet 'Attrition'
\begin{tabular}{lcccr}
\toprule
      &       & \multicolumn{3}{c}{Commitment arms} \\
\cmidrule{3-5}      & Control & Forced & Choice & \multicolumn{1}{c}{p-value} \\
\midrule
\midrule
Take-up & 0.97  & 0.96  & 0.97  & 0.53 \\
      & (0.01) & (0.01) & (0.01) &  \\
\midrule
Number of borrowers & 20.8  & 22.2  & 25.4  & 0.17 \\
      & (3.29) & (3.9) & (4.89) &  \\
\rowcolor[rgb]{ .949,  .949,  .949} \multicolumn{1}{r}{median} & 19    & 20    & 21    & 0.5 \\
\midrule
Number of pawns/borrower & 1.4   & 1.4   & 1.4   & 0.43 \\
      & (0.08) & (0.04) & (0.05) &  \\
\rowcolor[rgb]{ .949,  .949,  .949} \multicolumn{1}{r}{median} & 1.4   & 1.3   & 1.3   & 0.6 \\
\midrule
Number of pawns  & 31    & 31.3  & 37.2  & 0.24 \\
      & (5.8) & (5.6) & (7.9) &  \\
\rowcolor[rgb]{ .949,  .949,  .949} \multicolumn{1}{r}{median} & 27    & 28    & 30    & 0.46 \\
\midrule
Total borrowed & 2266.8 & 2094  & 2115.2 & 0.18 \\
      & (101.8) & (83.7) & (99.9) &  \\
\rowcolor[rgb]{ .949,  .949,  .949} \multicolumn{1}{r}{median} & 2154.3 & 2041  & 2047.5 & 0.65 \\
\midrule
Number of pawns/client & 1.4   & 1.3   & 1.4   & 0.1 \\
      & (0.07) & (0.04) & (0.04) &  \\
\rowcolor[rgb]{ .949,  .949,  .949} \multicolumn{1}{r}{median} & 1.4   & 1.3   & 1.3   & 0.71 \\
\midrule
Obs   & 85    & 81    & 94    &  \\
\bottomrule
\bottomrule
\end{tabular}%