### 2025-04-25 Restud submission/donde_restud-FIRST-DRAFT.tex

\documentclass[12pt, a4paper]{article}

\input{preamble.tex}
%%% HELPER CODE FOR DEALING WITH EXTERNAL REFERENCES
\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{
    \externaldocument{#1}
    \addFileDependency{#1.tex}
    \addFileDependency{#1.aux}
}

\newgeometry{margin=1in}
\linespread{1.25}
\begin{document}

\title{Structured Payment in Pawnshop Borrowing\thanks{We want to thank Jose Maria Barrero, Andrei Gomberg, Emilio Gutierrez, Anett John, David Laibson, Aprajit Mahajan, Matt Rabin, Mauricio Romero, Charlie Sprenger, Chris Udry, and seminar participants at ITAM, USC, MSU, UC Davis,  UCSD, and USC who provided valuable feedback. Ricardo Olivares, Gerardo Melendez, and Alonso de Gortari provided excellent research assistance and Erick Molina helped with formatting.  Research assistance was financed through faculty grants at ITAM. Our research partner had no say in the design or results of the experiment.}}
\author{Craig McIntosh \and Isaac Meza \and Joyce Sadka \and Enrique Seira \and Francis J.\ DiTraglia   \thanks{Seira:  MSU, \url{enrique.seira@gmail.com} (corresponding author); McIntosh:  University of California San Diego, \url{ctmcintosh@ucsd.edu}; Meza: Harvard University, \url{isaacmezalopez@g.harvard.edu}; Sadka: ITAM, \url{jsadka@itam.mx}; DiTraglia: Oxford, \url{francis.ditraglia@economics.ox.ac.uk}} }
\date{This draft:  \today \\[2 cm]}

%\vspace{.5in}

\maketitle
\vspace{-0.75in}

\begin{abstract}
We explore paternalism in the understudied context of high-default pawn lending. We use a novel one-stage, three-armed RCT, including forced and voluntary treatment arms, to point-identify the effects of structured repayment contracts on both the treated and the untreated simultaneously. We find that forcing borrowers into commitment-style contracts decreases the financial cost of pawn loans by 22\% and increases the likelihood of recovering their pawn by 15\%. However, only 11\% of borrowers choose such contracts. We find that borrowers who would not choose regular payments would benefit substantially from getting such contracts, as much as those freely choosing them (no selection on gains).  A paternalistic policy of universal forced commitment archives similar benefits to personalized targeting. 
\end{abstract}

\vspace{.8in}
\noindent \textbf{Keywords: } Paternalism, choice, treatment on the untreated, payment frequency, overconfidence. \textbf{JEL codes:} G41, C93, O16, G21

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Main text entry area:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section{Introduction}

Pawnshop lending, where borrowers provide gold and other easily sold items to receive overcolleralized loans, are one of the most prevalent forms of credit in the world. loans are one of the oldest and most common forms of borrowing \citep{carter2012pawnshops}. 
In Mexico the scale of pawn lending rivals that of microfinance: during the past three years a single large lender in the country made over 4 million loans to more than a million clients. Compare this to a total of 2.3 million microfinance clients in all of Mexico during 2009 \citep{Pedroza:2010}. Pawn borrowers are not only numerous, they are also economically vulnerable. The vast majority of pawn loans are made to people with relatively low levels of education in times of significant stress.\footnote{A large literature shows stress impairs cognitive function, e.g. \cite{StressReview}.} Despite its prevalence and at-risk clientele, however, pawn lending has received little attention in the literature compared to other financial services.

Financial institutions often use built-in commitment mechanisms like fixed repayment streams to lower default rates and increase their own profits. At the same time, they frequently ``shroud'' these attributes by embedding them as non-negotiable features of the loan contract, presumably because borrowers undervalue commitment and would not demand it, given the choice. \cite{Laibson2018} refers to this practice as ``private paternalism''. When the incentives of firm and client align, as they do in \cite{Laibson2018}, private paternalism can benefit both parties. Pawn lending, however, turns this idea on its head: since the value of the collateral--the pawn--exceeds that of the loan, lenders stand to gain the most when borrowers default.\footnote{An industry standard contract in Mexico City lends 70\% of the value of gold collateral at an interest rate of 7\% for three months. Because gold is highly liquid, the profit from a borrower who defaults is 30\% of the collateral value compared to 15.8\% in interest if a loan is paid in full on the last day.}
Consistent with this idea, and in stark contrast to microfinance, pawn lenders in Mexico City neither send payment reminders to borrowers nor provide commitment features in their contracts. Default is common.\footnote{The default rate in the \emph{status quo} arm of our Mexico City experiment is 44\%.}

This paper presents results from a randomized experiment that answers three key questions. First, does the absence of commitment-style contracts cause default in pawn lending? If so, to what extent does this benefit lenders? Second, if a commitment contract were available would pawn borrowers choose it? Even more importantly, would the \emph{right} borrowers choose it, i.e.\ those who would benefit the most? Third, is there a case for \emph{paternalism}, namely for mandating commitment in pawn lending? If so, should commitment be mandated for all borrowers or can it be effectively targeted? The question of choice versus paternalism is particularly salient in a setting with vulnerable borrowers. Yet comparing voluntary versus paternalistic policies is challenging. Even if commitment lowers financial costs on average, it may be ineffective or even harmful for borrowers who \emph{would not choose} to commit. To meet this challenge we introduce a novel three-armed design, the ``forcing-choice'' design, that allows us to shed light on these three questions in the important and under-studied context of pawn lending.

Our experiment covers just under 5,000 pawnshop clients across six of our partner lender's Mexico City branches. The control arm illustrates the costs of the \emph{status quo} contract: fully 44\% of control borrowers default, losing their pawn along with any payments made towards principal. Worse still, 29\% of those who default do so after making at least one payment towards recovery, a payment that they forfeit upon default. The ``forced commitment'' arm requires borrowers to make three monthly payments, with each monthly payment including the accrued interest at that time as well as a nominal fee of 2\% of that month's payment if the payment is delinquent.  Finally, the ``commitment choice'' arm offers borrowers the chance to opt in to the same structured monthly contract that is mandated for borrowers in the forced commitment arm.\footnote{When describing this new contract below, we use the terms ``commitment'' and ``structured payment'' interchangeably.}

Because there is no selection into or out of the control (status quo) and forced commitment arms in our experiment, the forcing-choice design can be viewed as a \emph{juxtaposition} of three more familiar experimental designs.\footnote{See section \ref{sec:integrity} for a discussion of how we can be sure that there was no selection into or out of the treatment and forced commitment arms.} The first is a standard two-armed RCT with perfect compliance that point identifies the average treatment effect (ATE) of commitment. The other two are a pair of randomized encouragement designs, each with one-sided non-compliance: one point identifies the effect of commitment for borrowers who would \emph{voluntarily} choose commitment (treatment on the treated: TOT), while the other point identifies the effect of commitment for borrowers who would \emph{not} (treatment on the untreated: TUT). 
Fully evaluating the case for paternalism versus borrower choice requires policymakers to learn not only whether those who demand commitment benefit from it (TOT), but also whether those who \emph{do not} are better off without it (TUT).

Our ATE results show that commitment is strongly effective in lowering financial costs and preventing default in pawnshop lending: the average borrower in the forced arm pays financing costs inclusive of fees that are 22\% lower than the control, and faces a probability of default that is 6.6 percentage points lower (15\% of the mean). In terms of Annual Percentage Rates (APR), the financial cost of borrowing falls by 11 percentage points per loan (19\% of the mean). In short, structured commitment saves borrowers money by charging them fees. The mechanism by which our monthly payment contract achieves these cost savings is intriguing: not only does it speed up payments, it also generates an early \emph{bifurcation} of borrowers into those that will recover their pawns and those that will not. The former are induced to pay faster, saving on interest; the latter pay less towards loans that ultimately default, hence losing less money when they do. While we find that borrowers who experience structured payment contracts are 6 percentage points more likely borrow again from the same lender, a back-of-the-envelope calculation suggests that the commitment contract nevertheless lowers lender profits by \hl{11\% per borrower}. It is thus unsurprising that commitment contracts were unavailable in Mexico city pawnshops outside of our experiment: lenders profit from default.\footnote{Unlike most pawn lenders, our partner operates as a non-profit, donating all of their earnings to philanthropic causes. They approached us and requested help evaluating the causal effect of commitment.}

Given the limited measures that are available to us, we refrain from making definitive welfare statements in this paper, focusing instead on the cost of loan servicing, an important outcome in credit markets.\footnote{\hl{Enrique will find citations for a defensive footnote that goes here, pointing out that lots of papers use this as an outcome measure.}} Nevertheless, we explore the sensitivity of our findings to broader notions of cost. For example, our results are qualitatively robust to deducting transport costs of visiting the branch to make interim payments along with a day of lost wages for each visit. They are also robust to using borrowers' \emph{subjective} values of their pawns rather than the appraised value and to adjusting for lost liquidity from requiring monthly payments.

Despite the large financial cost savings of structured repayment, only 11\% of borrowers in the choice arm opt into the commitment contract. 
Could it be that borrowers who do not choose commitment are those who simply don't need it? 
To answer this question, we carry out a detailed analysis of treatment effect heterogeneity.
We begin by bounding the distribution of individual treatment effects using the marginal outcome distributions from the forced commitment and control arms, following \cite{fan2010sharp}. This approach imposes no assumptions beyond the experimental randomization. We find that \emph{at least} 23\% of borrowers benefit from the structured payment contract. This implies that there must be many borrowers in the choice arm who did not demand this feature despite their \emph{individual treatment effect} from (forced) commitment being positive.

Next, we impose a pair of exclusion restrictions positing that the effect of a given contract does not depend on \emph{how} borrowers obtain it. In other words, we assume that choosing a contract results in the same potential outcome as being assigned that contract. This is a relatively common if often implicit assumption in causal inference.\footnote{Papers that use variation in compulsory schooling laws to identify the returns to schooling, for example, typically interpret their results as the causal effect of additional \emph{education} rather than additional \emph{forced education}. \cite{chamberlain2011bayesian} uses a closely related assumption to develop a theory of optimal treatment choice.} 
It also has testable implications that we fail to reject in our empirical context.\footnote{See\autoref{sec:testing_exclusion} \hl{for details--Frank will move some of this into the body of the text.}}
Under these restrictions, the forcing-choice design point-identifies the TOT and TUT, described above, along with the average selection on gains $\text{ASG} = \text{TOT} - \text{TUT}$, a measure of the empirical relevance of Roy-type selection.
Our estimated TUT effect on financial cost savings is large: \$192 pesos, equivalent to a 10.6 percentage point savings in APR. On average, the borrowers who would \emph{not} choose commitment, would have faced substantially lower financial costs if they had. 
Finally, we combine our experimental treatment and outcome data with survey responses collected for a subset of borrowers to estimate conditional average treatment effects, both TUTs and ATEs, using the Causal Random Forest algorithm of \cite{atheygrf}. We estimate positive conditional average TUT effects for 93\% of the borrowers who did \textit{not} choose the commitment contract. In short, it is extremely difficult to find identifiable groups of borrowers who incur higher financial costs under a commitment contract, even when restricting attention to those who would not choose it voluntarily.

So why do borrowers seem to be leaving money on the table? 
While a simple measure of time inconsistency does not explain the large and positive TUT effect, we find that borrowers systematically overestimate the probability that they will recover their pawn: the average subjective probability of recovery is 92\% compared to a realized probability of 44\%.
Indeed, among borrowers who choose not to commit, we estimate the largest gains from commitment for those who overestimate their probability of recovery \emph{the most}.
This suggests that overconfidence may reduce the demand for commitment.

According to our conditional TUT results discussed above, mandating commitment contracts for all borrowers runs the risk of increasing financial costs for 7\% of the untreated. This raises the question of whether it is possible to assign commitment contracts in a \emph{targeted} way to only those borrowers who are likely to benefit from them. To answer this question, we revisit the causal forest results described above for the subset of borrowers who completed our baseline survey. Importantly, some variables from our survey that predict treatment effect heterogeneity cannot be used for targeting in practice. For example, while overconfident borrowers are more likely to default, targeting based on subjective probability of recovery may incentivize borrowers to change their reports, causing the predictive relationship to break down. Accordingly, our exercise restricts attention to survey variables that would be hard to manipulate (``narrow'' targeting). With this restriction, all of the targeting rules that we explore have relatively weak predictive power. Our best-performing targeting rule only lower the overall mis-targeting rate from 9.7\%--the baseline when all borrowers are assigned to commitment--to 9.5\%.

Our paper makes a number of contributions to the literature. First, we provide a detailed picture of pawn lending, a widespread and important financial service that, despite its prevalence and impact on economically vulnerable borrowers, has received little scholarly attention. While an older literature considers the exploitative potential of over-collateralization and underpriced collateral \citep{basu1984implicit}, the behavioral implications of such contracts remain largely unexplored. To our knowledge, ours is the first behavioral paper to focus on pawn lending specifically.

Second, we contribute to the relatively small literature on paternalism and selection-on-gains. In contrast to \cite{Laibson2018}, who studies private paternalism from a theoretical perspective, we show empirically that pawn borrowers would benefit, and lenders would suffer, from a paternalistic policy that mandates commitment in the form of structured repayment contracts. Unlike \cite{Sprenger}, who shows that individuals with the most time-inconsistent preferences are the least likely to demand commitment, we directly identify the treatment effect on the untreated (TUT). This obviates the need to first elicit preferences before testing for negative selection. Furthermore, whereas \cite{Walters} combines a distance-based instrument with structural modeling assumptions to show that students who select into more effective schools have smaller treatment effects ($\text{TOT} < \text{TUT}$), we identify the TUT and TOT effects without the need for a structural model. Instead, we rely on relatively weak exclusion restrictions whose testable implications we fail to reject.

Third, our paper makes a methodological contribution by introducing the ``forcing-choice'' design, a novel three-armed experiment that point identifies treatment effects for choosers (TOT) and non-choosers (TUT) under minimal assumptions.\footnote{A companion STATA package provides simple, regression-based estimators for each of the causal parameters that are identified under this design, along with appropriate cluster-robust standard errors. See \autoref{sec:estimation_inference} for more details.}  
To our knowledge, this is the first paper to simultaneously point identify both quantities in the same experiment without the need for a structural model.\footnote{The forcing-choice design also point identifies the average selection-on-gains $\text{ASG} = \text{TOT} - \text{TUT}$, the average selection bias (ASB)--the difference in untreated potential outcomes for choosers versus non-choosers--and the average selection on levels (ASL)--the analogous comparison for treated potential outcomes. See Section \ref{sec:randchoice}.}
Our approach provides a valuable alternative to existing approaches for studying the relationship between treatment effects and treatment take-up. While the Marginal Treatment Effect (MTE) approach can be applied to either experimental or observational data provided that an instrumental variable is available, applying it requires additional modeling assumptions unless the instrument has a rich support set \citep{mogstad2018using}. The Becker-Degroot-Marschak (BDM) mechanism provides a theoretically attractive alternative that does not depend on structural modeling assumptions, but some recent studies have raised concerns about its reliability in practice \citep{bohm1997eliciting,banerji2014detection,mamadehussene2023reliability}.

The forcing-choice design differs from and complements other multi-armed designs from the literature.
As far as we are aware, our approach is the first to simultaneously identify both the TOT and TUT effects for within the same experiment, allowing us to study the empirical relevance of selection-on-gains.
The TUT effect, in particular, is crucial for considering the case for paternalism.
While \cite{fowlie2021default} likewise employ a three-armed experimental design in their study of the effect of electricity pricing, they instead identify \emph{two} TOT effects for different groups of ``treated'' households.
The well-known \cite{KarlanZinman2009} design, which randomly assigns interest rates in two stages to separate adverse selection from moral hazard, does not identify a TUT effect.
Like \cite{KarlanZinman2009}, \cite{Beaman2023} employ a two-stage randomization compared to the one-stage, three-armed approach in the forcing-choice design.
And while they identify a number of interesting causal parameters that shed light on the relationship between take-up and treatment effect heterogeneity, \cite{Beaman2023} do not identify a TUT effect or the average selection-on-gains.\footnote{In particular: the identify the ATE of cash grants, the TOT of loans, and the average effect of cash grants for those who would and would not accept a loan. While interesting, these last two effects are not TOT and TUT effects, because effects and compliance are defined with respect ti \emph{different} treatments: grants versus loans.}

Finally, we contribute to a growing literature on the effects of payment frequency in lending. Unlike previous experiments in microfinance markets \citep{Pande, barboni2023flexible, Field}, which find no benefits from regularized repayment environments, our experiment reveals significant positive effects in the context of pawn lending. The setting of these microfinance experiments differs from ours in an important way: they are performed on top of already highly structured microfinance contracts, and they involve borrower pools who may have selected into that type of lending precisely because it provides structure \citep{bauer2012behavioral}. This difference may explain why \cite{Pande} find almost no default in their control group, compared to 44\% in ours.
In addition to studying the effects of payment frequency, we examine both take-up and the efficacy of voluntary commitment mechanisms. A number of papers have found low demand for commitment as we do.\footnote{\cite{Ashraf}, \cite{Gine}, \cite{Ted}, \cite{Royer}, \cite{Sprenger}. Others have found more robust demand for commitment (\cite{Kremer},  \cite{Casaburi}, \cite{Alcohol}, \cite{AprajitP&P}, \cite{Pascaline}).} Unlike all of these papers, however, we separately point-identify and estimate the effects of commitment for borrowers who would and would not choose it.

The remainder of the paper is structured as follows:  Section \ref{context} provides context and defines our main outcome variables. Section \ref{Design} describes the experiment and data sources, and shows pre-treatment balance across arms. Section \ref{Experiment} provides the standard ITT analysis of the experiment, while Section \ref{Choice} shows how to identify, estimate and carry out inference for the TOT, TUT, ASG and related causal effects under the forcing-choice design. Section \ref{Paternalism} investigates why paternalism functions so well in this context and whether it can be more finely targeted and Section \ref{conclusion} concludes.

\section{Context and stylized facts} \label{context}

\subsection{Pawnshop borrowing}

Pawn loans involve individuals leaving valuable liquid assets, typically jewelry, as collateral in exchange for an immediate cash loan. Collateral is typically more valuable than the loan amount, allowing lenders to give the loan immediately without checking a borrower's credit history. This makes pawn loans a popular way to get cash to pay for emergencies. In fact, they are one of the most prevalent forms of borrowing. There are more than 11,000 pawn shops across the US, with 30 million clients and \$14 billion in yearly revenues.\footnote{See
\href{https://tinyurl.com/ybm56dpe}{here}, \href{https://tinyurl.com/y9zdcgws}{here}, and \href{https://tinyurl.com/y59ptdam}{here}.}
Our partner pawn lender alone served more than 1 million clients in the last 3 years with more than 4 million contracts. For comparison there were 2.3 million micro-finance clients across all lenders in Mexico in 2009 \citep{Pedroza:2010}.

Pawning is also one of the oldest forms of borrowing. Pawn lending existed in antiquity at least since the Roman Empire, and there are records of it in China about 1,500 years ago \citep{PawnShops}. In spite of the high prevalence and long history, pawnshop borrowing has not received much attention in the economics literature. The closest widely studied product is payday lending. In developing countries, however, payday lending is likely small compared to pawnshop lending; the latter is faster and requires less documentation, making it more accessible to informal sector workers who receive their salaries in cash. %According to \cite{Payday} to get a payday loan: ``All that a prospective borrower typically needs is a home address; a valid checking account; a driver’s license and Social Security number; a couple of pay stubs to verify employment; wages and pay dates; and minimum earnings of at least \$1,000 a month''. Although the author meant this list as an instance of low requirements, they would render virtually all poor households in the developing countries ineligible.

%Lender $P$ was interested in understanding why half of their clients lost their pawn. Default is high in this context  compared to many papers in the microfinance literature.\footnote{\cite{Pande} study of the effect of payment frequency in a context where the rate of default is close to 1\%.}

As with payday lending, pawnshop lending is controversial. Regulators have concerns with the sophistication of borrowers using it, speculating they may suffer from behavioral and cognitive deficiencies that lead to making sub-optimal choices, biases that are exacerbated by contract design.\footnote{The US congress has actually banned the payday lending industry from serving active military personnel, and some States in the US have imposed zoning restrictions, interest caps, and restrictions on serial borrowing as consumer protection measures against payday lending \citep{Payday}.} There is some evidence in support of this view for payday borrowers\footnote{\cite{Bertrand} write that ``Under the view that the people borrowing from payday lenders are making an informed, utility-maximizing choice given the constraints that they face, one would not expect additional information disclosure about the payday product to alter their borrowing behavior'', but to the contrary, they find that simply disclosing how financing costs add up reduced demand by 11\%. \cite{Meltzer} finds that payday loan access leads to increased difficulty paying mortgage, rent and utilities bills.} but none for the large pawn-lending industry. Our study reinforces the idea that a lack of sophistication may be an integral part of the way that standard pawn contracts are designed and structured by lenders.

\subsection{Pawning Logistics and Contracts}

To study this market, we partnered with one of the largest pawn shops in Mexico, an institution with more than one hundred branches spanning multiple states in Mexico. This lender (whom we refer to as `Lender P') has a simple and typical business model.

%\hfill \newline
%\noindent \textbf{Appraising and lending.} 
\paragraph*{Appraising and Lending} Lender P takes gold jewelry as collateral in exchange for a fraction of the value of the piece, in cash. No other collateral and no credit history checks are needed. The transaction takes less than 10 minutes and is conducted at the branch in person between the client and the appraiser (i.e. a teller).
%, see Figure \ref{PawnshopPicture}). 
The appraiser weighs the gold piece and runs tests on its purity. Based on these she assigns a gold value to the piece, stores it as collateral, and gives 70\% of the gold value of the piece in cash to the client. The borrower signs a 2-page contract with the conditions of the loan and leaves with the cash. Appraisers were paid a flat monthly wage.

%\hfill \newline
%\noindent \textbf{Contract.} 
\paragraph*{Contract} Lender P had only one type of contract, henceforth the \textit{status quo} contract. It stipulated that the interest rate was 7\% \textit{per month} compounded daily on the outstanding amount of the loan. The loan had a 90 days term with 15 days' grace period. The client could make payments at the branch at any time with no penalty for pre-payment. Under this status quo contract, there are no payment reminders or any other kind of interim contact between the lender and the borrower. If the client returns to pay the principal plus the accumulated interest within 105 days, she recovers her pawn, otherwise the pawnbroker keeps the piece \textit{and} any payments already made. Before the contract expires, the client had the right to renew for another 3 months by going to the pawnshop, paying the accumulated interest, and signing a new contract with exactly the same terms and the same piece as the original contract (38\% of borrowers renew at least once with a given pawn). This contract is standard in the industry.  Pawnshops make money in three ways: by reselling the jewelry left as collateral on defaulted loans, by charging interest on non-defaulted loans, and by keeping the payments made on defaulted loans.

\subsection{Stylized facts}

%\hfill \newline
%\noindent \textbf{Borrowers.} 
\paragraph*{Borrowers.} Borrowers are not completely inexperienced: 87\% of clients in our survey report they have pawned before. %These clients have little or no access to other types of loans and they value the convenience of pawn borrowing.
However they are economically vulnerable:  30\% of them could not pay either water, electricity \& gas or rent in the past 6 months. In terms of triggers and uses of the loan, 89\% said they are pawning because of an emergency, and only 12\% stated it was to be used in a ``non-urgent expense''.  When asked why they are pawning this piece, 5\% responded ``lost a family member'', ``a medical emergency'' (11\%), or ``an urgent expense'' (72\%). Only 32\% of borrowers have completed college, 73\% of are women, with an average age of 43 years.

%\hfill \newline
%\noindent \textbf{Many borrowers lose their pawn.} 
\paragraph*{Many borrowers lose their pawn.}
Our context is also one with high borrower default: 43\% of clients lose their pawn in a time span of 230 days from the date of pawning. One potential explanation for high default is that clients are really just knowingly selling their gold piece through a pawn contract on which they intend to default. This appears unlikely for several reasons: clients can easily sell the gold and obtain a higher amount of instant cash at gold-buying stores located close to almost all our pawnshop branches;
%(see Figure \ref{GoldBuyers}) 
the reported subjective value of the pawn is larger than the appraised value for 83\% of clients (this also means it is inefficient that they lose their pawns). The default rates for the experimental period and branches is similar to that observed for all branches in subsequent periods (see Figure \ref{external_validity}).

\paragraph*{Gold buying shops.} Gold purchasing shops locate close to our Lender's branches. Branches have an average of \hl{XX} of these shops.  This fact is important given the high default rates which could indicate that borrowers are ex ante intending to `sell' their pawn.  Given that they are receiving 70\% of the value of the pawn next door to shops that would give them 100\% of this value, we think this is prima facie unlikely.

\paragraph*{Paying without recovering.} Among those that lose their pawn, 29\% paid a positive amount towards its recovery, and on average paid 42\% of the value of their loan (see Figure \ref{proxy_naive} in Appendix) --- this can only be rationalized if they expected to recover their pawn.

\paragraph*{Overconfidence.} We implemented a baseline survey to more than 6,000 respondents. In it, we asked borrowers for their subjective probability of recovering the pawn. The average recovery probability is 92\%. This contrasts sharply with the actual recovery rate for these same borrowers, 44\%. Overall 72\% of borrowers report a 100\% probability of repaying their loan.    %Note that high default could be detrimental from the lender's point of view, since it may reduce the likelihood that the client becomes a return customer. 
%Lender P was in fact explicitly interested in partnering with us to investigate how to reduce loan default, customer satisfaction, and repeat borrowing.

\subsection{Measuring Borrowers Financial Costs} 
\label{costs}

Borrowers' financial costs are composed of two main categories: the cost of losing their collateral, and the interest and fees incurred during the life of the loan. For each given loan we observe if the client lost her pawn, $\mathds{1}(\text{Default}_i)$. If a loan has been rolled over and is still outstanding, we consider it to be non-defaulted.  This approach is conservative in our context (biases treatment effects towards zero), as we show in detail in Section \ref{Experiment}. In our data 13\% of experimental loans are ongoing (i.e. censored) when the data period ends. Regarding interest, our administrative data classifies payments made into three types according to their payment allocation rules: payments to principal $P^C$, payments on generated interests $P^I$, and payments on penalty fees $P^F$. We observe each and every payment made under each category, its amount and date.

We define a borrower's financial cost as the total monetary outflow---in cash or pawn value---from the borrower to the lender. This includes all payments the borrower made toward interest and fees, but also the net difference between the appraised value of the pawn and the loan amount (Value-Loan) in the event of default. When there is no default the borrower gets her pawn back and there is no loss of value for the borrower. Payments towards capital are considered a cost only when the borrower defaults, as she does not get reimbursed for these. Note however that when she does not default payments to capital are not an actual outflow, as they sum up to the value of the loan the lender disbursed in the first place. The formula for financial cost for person $i$ is thus as follows: %\footnote{We alternatively defined financial cost as $\text{Financial Cost}_i = \sum_t P^I_{it} + \sum_t P^F_{it} + \sum_t P^C_{it} + \mathds{1}(\text{Default}_i) \times (\text{Value}_i)-\text{Loan}_i$, and obtained similar results.}:
\begin{align*}
    \text{Financial Cost}_i =&  \sum_t P^I_{it} +\sum_t P^F_{it}  
     + \mathds{1}(\text{Default}_i) \times \left(\text{Value}_i-\text{Loan}_i + \sum_t P^c_{it}\right)
\end{align*}
where $t$ indexes days, and $\mathds{1}(\text{Default}_i)$ is an indicator function for defaulting. Because the period of the loan is only 90 days we do not apply discounting in calculating costs.  In robustness checks reported below we show that our results are virtually unchanged when applying a wide range of time discounting factors.

We consider the above to be an accurate measure of financial cost in pesos. However, we also report results incorporating two non-financial costs: (i) using the subjective value of the pawn reported by the borrower in place of its appraised gold value, and (ii) adding a measure of travel expenses and the opportunity cost of time, as clients have to go to the branch in order to make payments.

As a second measure of cost we calculate the Annual Percentage Rate (APR) in order to express the cost as a percentage of the loan, per year, inclusive of default costs. The standard definition is given in the following formula: 
$$(\text{APR})_i =\left( 1 + \frac{\frac{\text{Financial Cost}_i}{\text{Loan}_i}}{\text{loan term}_i}\right)^{\text{loan term}_i}-1 $$

%We calculate an APR of 218\% on average for the control group.%\footnote{The Annual Percentage Rate (for pawn $j$) is calculated as the internal rate of return $i$ such that $\sum_t \frac{P_{jt}}{(1+i)^t} + \frac{\text{I}(Default \: Pawn_j) \times Pawn \: Value_j}{(1+i)^T} - Pawn \: Value_j = 0$, where $T$ is the date the pawn was lost, and $P_{jt} =P_{jt}^c+P_{jt}^f+P_{jt}^i $ is the sum of all payments.}

\vspace{.2in}
\section{Experiment and Data} \label{Design}

%\hfill \newline
\subsection{Treatment arms and randomization}

%\hfill \newline
%\noindent \textbf{The Commitment Contract.} 
\paragraph*{The Commitment contract} For the purpose of the experiment we designed a new contract that is identical to the status quo contract except that, informed by the design of micro-lending contracts, it enhances the regularity and salience of payments as a way to encourage repayment \citep{morduch1999microfinance, bauer2012behavioral}.  It has the same interest rate (7\% \textit{per month}) which accumulates daily on outstanding debt, the same loan size/collateral ratio (70\%), and the same loan term (90 days, and a grace period of 15 days). Borrowers' gold pawns are appraised in the same way by the same appraisers under both the new and status quo contracts. The commitment contract however requires the client to make regular monthly payments for the duration of the contract, with the principal and interest payments split evenly across the three months of the contract (day 30, 60 and 90 after loan disbursement). The importance of this monthly payment was made salient in the contract and payment receipts, and by the levying of a nominal fee (2\% of minimum due) on individuals who fell behind in their payments. %\textcolor{red}{The idea that the fee itself is not driving the treatment is reinforced later in the paper where we show financial benefits many times larger than the fee, as well as treatment effects even on those who would fail to recover pawns and hence not pay the fee.} 
The fee was modest and intended to make the payment deadlines salient. As a benchmark, the transportation cost to visit the branch to make a payment is comparable to the fee, on average.
%\footnote{%The lender was worried that if the fee was large, forgetting to pay could make the pawn unrecoverable. 
%In contrast with \cite{John} we don't let the clients choose the size of the fee. \cite{John} shows that doing this results in costly choices by (partially) naive present biased consumers. Our fee is much smaller than the median in her experiment. %We also experimented with a contract where there was no pecuniary fee for paying late but where we clients made a personal promise to pay monthly.}

To elicit demand for the monthly payment contract, we include an arm that allows borrowers to opt into this contract if they choose. The existence of both a non-optional ``forcing'' arm, and a choice arm in our design is key to estimating a battery of treatment effects under fairly mild assumptions. We next describe the three experimental arms in more detail.

%In order to explore lower-cost forms of structure, we also include arms that feature `soft' commitment, asking or requiring individuals to follow the more regular payment structure of the commitment contract but then providing no reminders or repercussions for failing to do so.  A substantial literature documents that people are averse to to breaking their promises, and are willing to incur in monetary losses in trying to keep them. This provides a potentially very attractive way to drive better repayment behavior at low pecuniary cost to clients.  Numerous lab experiments and at least one field experiment on financial services show that such soft commitment can work \citep{Craig}.\footnote{Grameen Bank for instance makes clients recite a promise at each center weekly meeting.The exact promise repeated by clients live is this: ``\textit{We pledge to attend regularly the weekly Center meetings, to utilize our loans for the purpose approved, to save and pay our installments weekly, to use our increased incomes for the benefit of our families, to ensure that other members of our group and Center do likewise and to take collective responsibility if they do not.''}}  The soft treatments mimic the commitment arms as closely as possible without imposing external remainders or pecuniary consequences for failing to meet terms.

%\hfill \newline
%\noindent \textbf{Treatment Arms.} 
\paragraph*{Treatment Arms} 
Treatments were randomized at the branch-day level. Each day a computer randomly assigned which types of contracts were on offer that day in the branch, and the IT system would only offer these.  We have 3 different experimental arms\footnote{The experiment included other independent arms that involved no fee penalties and did not emphasize the structure of payments. These are being analyzed in a separate paper.}

\begin{enumerate}
    \item \textit{Control} arm: consisted of branch-days offering the status quo contract described in Section \ref{context}, and only this contract. 
    \item \textit{Forced structured payments} arm: consisted of branch-days requiring all borrowers to use the Commitment contract described above.  
    \item \textit{Structured payments Choice} arm: consisted of branch-days offering the client \textit{a choice} between the structured payments contract, and the status quo contract.
    %\item \textit{Forced Soft Commitment} arm: consisted of branch-days in which borrowers were asked to sign a non-binding written commitment to pay monthly as in the Commitment contract, but no reminders or fees were imposed if they did not.\footnote{The client was made to sign a paper which said ``I promise to pay every month the corresponding sum of \_\_\_\_\_\_, on the dates \_\_\_\_\_\_, \_\_\_\_\_\_, and \_\_\_\_\_\_. This is \underline{not} a legal document and cannot be used in courts. It is just a \textit{personal promise}. If I do not comply I will not have kept my word''. After signing, the promise was read to the client by the appraiser.}%\footnote{Our setting does not allow us to separate if the client is making a promise to the bank teller or whether she is mentally interpreting this as an internal promise to herself as a goal. We thank Anett John for pointing this out.}
    %\item \textit{Soft Commitment Choice} arm: consisted of branch-days on which the client was given \textit{a choice} between the Soft Commitment and the status quo contract.
\end{enumerate}

We did not allocate an equal number of days across arms, since we were interested in having more power in some of them. The number of branch days allocated to each were 84 to control, 80 to forced structured payments, and 93 to choice. See Figure \ref{exp_description} for a CONSORT-style diagram of the study design and recruitment.

\vspace{.2in}

\begin{figure}[htbp]    
 \caption{Experiment description}
\begin{center}
\includegraphics[width=0.9\textwidth]{Figuras/consort.pdf}
  \end{center}
     \label{exp_description}
%\textit{Do file: }  \texttt{consort.do}
\end{figure}

%\hfill \newline
%\noindent \textbf{Randomization.}  
\paragraph*{Randomization} We implemented the experiment in six branches of Lender $P$ beginning on September 9, 2012. The branches were selected by Lender $P$ to be dispersed across Mexico City and have varying sizes. In four of them the experiment ran for 102 days, and in 2 of them we ran it for a shorter time to economize on data collection costs once we realized we would not be constrained by sample size. %Thus we eliminated {the smallest} ones. 
Branches are more than 5 km apart from each other, and there is no substitution among them; none of the consumers appear in more than one of our branches.

Branch personnel did not know which treatment would be assigned to each day and were blind to the objective of the intervention. They were told that there were 3 different ``types of contract-days'', that the system chose randomly for any given date, and that it could happen for instance that two or more consecutive dates had the same contract. They were also told that this way of operating was in place in several of Lender P's branches (they did not know which ones), and that it would be in place for several months. Randomizing at the day level limits the problem of contamination arising from clients realizing that other clients get different contracts than theirs. It also limits potential manipulation by appraisers, who in the presence of individual-level randomization could potentially pick their preferred customer from the line or tell them to wait until their desired contract shows up on the screen. Intra-branch day correlation on the probability of default (ICC) is small, at {0.05}, so we lose little power vis-a-vis individual-level randomization.

Some clients pawned more than one time during the duration of the experiment, with 14\% pawning 2 times and 8\% more than 2 times. To have a clean comparison we are considering only the first pawn conducted during the experimental window. It is also the case that 30\% of those first pawns involve more than 1 loan, as 2 or more pieces of gold were submitted. We treat each of them as separate loans. In the appendix we show that our results are robust to this analysis choice.

%\hfill \newline
%\noindent \textbf{Timeline.} 
\paragraph*{Timeline} Figure \ref{exp_description} shows the experimental timeline along with the length of time for which we observe payments. For loans made in the first week of the experiment, we observe up to 338 subsequent days of loan information; for loans made in the last week we observe up to 235 days. Figure \ref{exp_description} also illustrates the number branch-days per arm, the number of loans, and the number of surveys. %Not every contract has a survey for several reasons, we have survey information for \hl{XXX\%} of them. There were two reasons: \hl{XXX}.

%\hfill \newline
%\noindent \textbf{Explanation.} 
\paragraph*{Explaining the Contracts}
We made sure clients understood the contract terms.  First, we had full-time enumerators explaining contract terms to clients. The explanation took about 3-5 minutes and continued until the client said she understood the contract terms. Enumerators then asked clients to explain the contract back to them before correcting any misunderstandings. Second, the appraiser gave clients the ``Contract Terms Summary'' and read it out loud to them before after their piece had been appraised but before they signed the contract. %The sheet clearly indicates that this contract is a monthly payment one (numeral 1), that there is a penalty of 2\% for paying late (numeral 2), and the 3 payment dates (numeral 3). 
%Finally, the bottom of the Figure \ref{PaperSlip} shows the paper slip we used for the promise arms. The clients had to put their name on a slip of paper where they stated they promised to pay monthly.
We are confident the overwhelming majority of clients understood the contracts and that those in the choice arm made informed choices.
%\footnote{That we can systematically predict demand based on consumer characteristics and measured beliefs suggest that take up is not random.}

\subsection{Data}

%\hfill \newline
%\noindent \textbf{Administrative data.} 
\paragraph*{Administrative Data}
The study exploits two types of data: administrative data from the lender, and a short survey that we implemented. The administrative data contains a unique identifier for each client, an identifier for the piece she is pawning, and the transactions relating to that piece. These transactions include the value of the item as assessed by the appraiser, the amount of money loaned (70\% of the item's value), the date of the pawn transaction, and the type of contract for that pawn: structured payments or status quo. Within the period of the loan, we followed each transaction related to that piece in the administrative data: when payments were made and for what amounts, whether there was default (i.e.\ the client lost her pawn), and whether any late-payment fees were imposed. After the experimental loan, we are able to track subsequent behavior and to see whether that borrower took a subsequent loan.  We have this information for all the pawns that occurred in the experiment's 6 branches between August 2, 2012 and August 13, 2013. This includes all the pawns that took place during our experiment along with those that one month before and eight months after our experiment. Figure \ref{exp_description} shows the design and timing of the experiment, along with the sample sizes in each arm. The experiment comprises 8,519 pawns while our administrative data covers a total of 26,180 pawns.

%\hfill \newline
%\noindent \textbf{Survey data.} 
\paragraph*{Survey Data} An additional team of enumerators was stationed at the entrance of each branch; they asked clients to complete a 5-minute survey before going to the teller window to appraise their piece and \textit{before} they learned which contracts would be available on that day. The survey allows us to measure loan takeup since we know whether people entered the branch, completed the survey, and then, after learning the contract terms, decided not to pawn. As we will see, this is extremely infrequent; only about 2\% of borrowers did that. This is not very surprising, borrowers come determined to borrow, many of them to solve an emergency, and are not very sensitive to contract terms (that may explain why they accept high interest rates and over-collateralization in the first place). The survey was intentionally brief. It measured demographics, proxies for income/wealth, education, present-biased preferences, experience pawning, if family or friends commonly asked for money, how time-consuming and costly it was to come to the branch, the subjective probability of recovering the piece that they intended to pawn, the subjective value of their piece in money terms (how much money they would sell it for), among others. We surveyed 7,210 clients. %, and our survey response rate was 78\% among clients who took loans. %\footnote{We also surveyed clients before and after the experiment, so our survey covers a larger sample that we are not using in this paper.}. 
Is it important to say that we do \textit{not} use the survey for our main results. We only use the survey in Section \ref{Paternalism} of the paper.

\subsection{No differential selection or imbalance}
\label{sec:integrity}

%\hfill \newline
%\noindent \textbf{Attrition.}

We were aware that the major threat for this study was selection into borrowing based on contract terms. This did \textit{not} happen. We show this in three different ways. First, out of those coming into the branch and filling out the baseline survey (at which time they did not yet know their treatment status), 97\% end up pawning. This rate is not only high but is virtually identical across treatment arms. Second, the number of loans per day is not different across arms, showing borrowers did not stop pawning when the forcing arm or choice arm was active.  Third, borrower and loan characteristics are the same across arms, showing that there is no selection on observables.\footnote{Note that is borrowers self-selected based on contract terms, the results would still reflect the overall impact of a treatment for the lender's portfolio. In our case, there is no selection, so we can study the causal effects on borrowers.}

An important implementation detail is that appraisers were paid a flat wage, independent of how many contracts or which types of contracts borrowers signed. They therefore had no incentive to influence borrower choice. Consistent with this, adding or removing appraiser fixed effects leaves results unchanged. 
%We should also recall that the lender's management team initiated this research collaboration and genuinely wanted to rigorously measure the effects of a structured payment contract. Our lender is a non-profit and donates all profits to philanthropy. They had no profit incentive to sway the results either way and could not do it anyway as researchers designed, implemented, and analyzed the data without intervention from the lender.

Table \ref{attrition_table_final} shows there are no differences in the number of loans, borrowers, or amount borrowed across arms using administrative data. Each row corresponds to a regression of the dependent variable on indicators for treatment assignment (status quo/control, forced commitment, choice). In the first row (``Take-up'') we use the sample of those who answered the baseline survey (6064 subjects) and check which of them pawned based on the administrative data. The dependent variable indicates whether the person pawned. Because this survey was implemented at the entrance of the branch before they knew their treatment assignment, it is a good measure of take-up. The first result is that the overwhelming majority (97\%) take the loan. Second, the same fraction of borrowers take up the loan across treatment arms. This is strong evidence that borrowers did not select deferentially according to the structured payments contractual feature. The remaining rows use only administrative data. For each branch day, we calculate the number of borrowers pawning, the number of pawns used, pawns per borrower, and the total amount borrowed and regress each of those on treatment arms. We again cannot reject that any of those measures of demand are the same across arms.  We also report medians to reduce the influence of outliers.

\begin{table}[H]
\caption{No selection across arms}
\label{attrition_table_final}
\begin{center}
\footnotesize{\input{./Tables/Attrition.tex}}
\end{center}
\scriptsize{Each row in this table 
 corresponds to a regression at the branch-day level of each dependent variable against the experimental arms indicators (control, forced commitment, choice). The table reports the coefficients on each of these indicators, as well as the p-value of an F-test of the null hypothesis of equality of the three coefficients. We also report results using median regression. %Figure \ref{boxplot_attrition}, in the appendix shows a box-plot of the first two variables, showing the distribution across arms is balanced.
 }
%\textit{Do file: } \texttt{ss\_att.do}
\end{table}

%\paragraph*{Attrition} There are two main channels through which attrition could complicate the interpretation of our results. The first, and more serious, is the possibility that clients might change their pawning decisions in response to the treatment they encounter in a given branch on the day they enter the branch.  If this occurred it would introduce a self-selection dimension which would still reflect the overall impact of a treatment for the lender's portfolio but would no longer deliver \textit{ceteris paribus} effects of treatments on individual borrowers.  Narrative reports and the way the treatment was implemented make us believe that selection into treatment is unlikely.\footnote{Potential clients did not know that different days could have different contracts. If they asked, appraisers said that whatever was offered on that day was the only available contract for an undetermined length of time. Anecdotally, appraisers told us that they did not think refusals differed across arms, and our enumerators informed us that potential clients rarely left the branch without pawning. Lender P also never complained to us that our different treatments were hurting sales.}

%If the treatments had induced demand-side selection, we would expect to see that the number of pawns successfully conducted differ in a systematic way across arms. That is, if potential borrowers disliked being forced into a structured payments contract, we would expect a lower number of pawns on branch days where only the structured payments contract is available compared to control days. Table \ref{attrition_table} shows that this is not the case. There is no difference at all between the Control and Forced structured payments arm in terms of the number of pawns per branch-day. %Although we cannot reject equality across the three arms (p-value=0.21), the Choice arm appears to have a somewhat larger number of borrowers than the Control and Forced arms.\footnote{p-value=0.23 and 0.12 respectively.} This seems to be due to sampling variability. The differences across arms are smaller when we look at medians, or when we look at the number of borrowers pawning (some borrowers pawn more than one piece). Figure \ref{boxplot_attrition} shows that full distribution of loans per day is very similar across treatment arms. The only noticeable difference is that the choice arm has a few more positive outliers.\footnote{If we winzorize at the 90th percentile, the mean number of branch-day pawns becomes identical across arms.}
%The bottom panel of Table \ref{attrition_table} shows balance in a more focused way, given that the surveys were conducted prior to the revelation of treatment status. We find that in no arm did more than three percent of individuals who responded to the survey go on to refuse loans. That is, the overwhelming marjority of potential borrowers did not leave the branch after learning which contract was on offer. Moreover, the extremely small fraction that did leave is balanced across arms.  Therefore it appears that the treatments have not induced any endogenous shifts in the composition of borrowers. 
%This is consistent with the findings of Table \ref{SS}.

% To test for this, we examine the number of loans given per branch-day, with specific attention to whether the Forced Commitment arm led to a lower number of borrowers per day. The first row of Table \ref{attrition_table} shows there is no difference between the Control and Forced Commitment arm in terms of the number of pawns per branch-day.\footnote{The Choice arm appears to have a somewhat larger number of borrowers than the Control arm (p-value=0.06), but we cannot reject equality with the Forced Commitment arm (p-value=0.41), or across all arms (p-value=0.16).}  The second row makes the same point in a more focused way, given that the surveys were conducted prior to the revelation of treatment status.  In no arm did more than three percent of individuals who responded to the survey go on to refuse loans, and neither these rates nor the rate of survey response are different across arms.  Therefore it appears that the treatments have not induced any endogenous shifts in the composition of borrowers.

% A less critical form of attrition is differential refusal to answer the survey questions.  The survey was conducted before treatment status was revealed, and we observe loan outcomes regardless of whether the survey was conducted.  Our core experimental estimates do not use the survey data as covariates, but the analysis in Section \ref{Paternalism} is restricted to the subset of borrowers who answered at least some survey questions. The bottom row of Table \ref{attrition_table} shows that the survey response rate is broadly similar across arms (about 78 percent).

%We can also examine differences in the fraction of borrowers who do respond to the survey and then do not end up taking a loan subsequent to completing the survey, we find that the share is close to 100\% as is the fraction of those 97\% across arms (p-value=0.62).
%\footnote{We also tested if the number of pawns per day for the 6 branches that participated in the experiment was different 1 month before the experiment started or 1 month after the experiment ended, relative to the days the experiment was active. To this end we estimated the regression. We cluster the standard errors by branch.} $Pawns \: per \: day_{jt} = \alpha_j + \gamma f(t) + \beta_b \mathbbm{1}(t \in MB)_{t} +\beta_a \mathbbm{1}(t \in MA)_{t}$, where $\alpha_j$ are branch fixed effects, $f(t)$ is a third degree polynomial in time, $\beta_b$ measures a level effect in the number of pawns per branch a month before the experiment started, and $\beta_a$ a month after. We estimate no difference in the number of pawns, showing that people are not leaving the branch without pawning at larger rates during the experiment ($\beta_a=1.32$, $\beta_b=-0.65$, with p-values 0.11 and 0.83, respectively).  Table \ref{num_pawns_bal} shows the beta coefficients from these regressions, which are nowhere significant.}  Hence while survey response was not universal, there is no evidence that it was differential across arms.

%A more subtle form of sample selection could arise if the treatments induce borrowers to re-pawn in different ways, especially given that their treatment status on subsequent loan/days may not be the same as that initially assigned.  To address this issue our analysis uses only the first loan taken by each borrower during the experimental window.

%\subsection{Connection to the literature}
%Our study connects with two strands of the literature on microfinance. But also to a nascent one that uses RCTs to study to what extent people self-sort to treatments that have larger beneficial treatment effects for them. On the microfinance literature the first paper we could find that used an RCT to evaluate the causal effect of frequent payments in loans is \cite{Pande}. In a group lending rosca context, they note that most contracts involved frequent repayments --even weekly in many instances-- even when this increases transaction costs. They note that clients could benefit from ``the fiscal discipline afforded by the more rigid payments''. Frequency could provide a commitment device for clients, could foster a payment habit, or could generate more trust from social interactions among the group of borrowers. All these potential benefits apply in our context, except those relating to social interactions, as we work with individualized loans.

%\subsection{More on implementation} %\label{implementation}

%\hfill \newline
%\noindent \textbf{Balance.} 
%\paragraph*{Balance}

Table \ref{SS_final} presents summary statistics for the sample of actual borrowers across arms, showing that our randomization succeeded in achieving balance across the experimental arms. %Panel A uses administrative data for the universe of borrowers in each arm, and shows that loan balances and the days on the week on which individuals pawned are comparable across arms. The average loan size is \$2267 MXN (\$130 USD). Panel B of Table \ref{SS} reports summary statistics across arms from our survey data. 
%Among the 78\% of borrowers who completed our survey, 
73\% of surveyed borrrowers are women, the average age is 43 years, 66\% have completed at least a high school education, and 87\% have pawned before. Finally, borrowers' subjective probability of recovering their pawn is close to 92\%. The average subjective value they report for the items they pawn is 4084 MXN, much larger than the average appraised gold value of 3238 MXN. %While this could arise either from overconfidence in valuation or from undervaluation by the lender, in any case it is \textit{prima facie} evidence that loss of the pawn should be undesirable relative to the quantity of liquidity leveraged by the asset. 
Importantly, all borrower's characteristics are balanced across arms, again consistent with no differential selection across arms. Tables \ref{attrition_table_final} and Table \ref{SS_final}  tests are the standard experimental integrity analysis. Together they constitute strong evidence of no borrower selection across arms. This was reinforced by reports from our enumerators and appraisers that those who entered the branch were determined to borrow and that there were no complaints about what loan contracts were available. Note that this does not mean that they did not care about contract terms. It means just that the outside option of leaving the branch without a loan is unattractive.

\begin{table}[H]
\caption{Borrower's characteristics are balanced}
\label{SS_final}
\begin{center}
\resizebox{0.65\textwidth}{!}{
\footnotesize{\input{./Tables/SS.tex}}
}
\end{center}
\scriptsize {The table uses survey data to show balance across arms. Survey data is not used for the main results of the paper. It is only used in Section \ref{Paternalism}. Each row in this table corresponds to a regression, where the level of observation is a borrower that pawned which was surveyed.  The dependent variables are displayed in the first column. Each dependent variable is regressed against the experimental arms indicators (control, forced commitment, choice). The table reports the coefficients on each of these indicators, as well as the p-value an F-test of the null hypothesis of equality of the three coefficients. ``Subjective value'' of the pawn refers to how much would the client be willing to sell the pawn for in pesos, ``Trouble paying bills'' is an indicator for the borrower reporting having problems paying electricity, water and services in the last 6-months, present bias is constructed as in \cite{Ashraf}; ``Makes buget'' os an indicator for whether the household make expenses budget for the month ahead of time. The subjective probability of recovery was elicited a la Manski (from 0 to 100 what is the probability that you will recoup your pawn); pawned before is a dummy=1 if the client declares to have pawned before (although not necessarily with Lender $P$); age is age of the borrower in years, +High-school is a dummy that indicates if the client has completed high school. 
}
%\textit{Do file: } \texttt{ss\_balance.do}
\end{table}

\section{Average Treatment Effects} \label{Experiment}

We begin by estimating average treatment effects of assignment to the Forced structured payments and the Choice arms, relative to those assigned to the Control arm. As we explain below, only about 11\% of those in the choice arm chose the monthly payment option (Figure \ref{determinants_choose} shows coefficient plots for the characteristics that determine choosing structured payments in the Choice arm).

%\subsection{Average Treatment Effects}

%\hfill \newline
%\noindent \textbf{Specification.} 
\paragraph*{Specification} Table \ref{main_impact_table} presents estimates and standard errors from a standard pooled experimental regression 
\begin{equation} \label{basic_reg}
    y_{ij} = \alpha + \beta^F T_{i}^F + \beta^C T_{i}^C + \gamma X_{ij} + \epsilon_{ij}
\end{equation}
where $i$ indexes the borrower, $j$ indexes branch, $T_{i}^F$ and $T_{i}^C$ are indicator variables for receiving the Forced or Choice arms, $X_{ij}$ are branch and day-of-week fixed effects. Standard errors are clustered at the branch-day level, the unit of treatment assignment\footnote{A minority of borrowers pawned on more than one day during the experiment: 14\% pawned on two distinct days, and 8\% on three or more days. To avoid contamination from earlier treatments to which these individuals were exposed, we restrict our sample to each borrower's \emph{first visit}. Note that a borrower may pawn multiple items her first visit. We include these as separate observations. Because our standard errors are clustered at the branch-day level, they automatically account for any dependence in error terms arising from multiple pawns by the same borrower on her first visit.}.
Given full compliance in the Forced arm, the coefficient $\beta^F$ is the ATE of forced commitment while $\beta^C$ is the ITT of commitment in the Choice arm on the outcome variable $y_{ij}$.
%\footnote{Note that $\beta^C$ can be viewed in two ways: as the ATE of \emph{offering clients a choice} between the status quo and commitment contract or as the intent-to-treat effect of the commitment contract \emph{itself}. Here we adopt the former interpretation; below we use the latter to consider the relevance of selection-on-gains.}
Our two primary outcome variables are financial cost in pesos and Annual Percentage Rate (APR), as defined in Section \ref{costs}.

Results for these outcomes appear in columns (1) and (7) of Table \ref{main_impact_table}.
The remaining columns decompose the financial cost and APR outcomes into their components: interest payments (col 2), payment towards any fees incurred (col 3), payments toward the principal (col 4). Column 5 shows the value of lost pawn conditional on losing it.  In column 6 the dependent variable is a dummy indicating default.  Finally, column 7 rescales financial cost as a function of loan size to estimate causal effects on incurred APR.\footnote{As we explained above, loans can be extended for an additional 3 months by paying the interest owed and restarting the loan under the same treatment conditions. This means that some loans extend for more than 3 months. We consider the entire flow of cost for the duration of our sample.}

\begin{table}
\caption{Effects on Financial Cost}
\label{main_impact_table}
\begin{center}
\resizebox{1.0\textwidth}{!}{
\footnotesize{\input{./Tables/decomposition_main_te.tex}}
}
\end{center}
\scriptsize{This table shows the treatment effects for our core pecuniary outcomes. Each column is a different regression for different outcomes on an indicator for the forced and choice arms, following specification in equation \ref{basic_reg}. Columns (1) \& (7) analyze our core financial cost measures, while the rest of the columns decompose these into finer components.
A few borrowers take more than one loan on the first day they appear in an experimental branch. These are treated as different observations. Additional results, available upon request, show that our results are robust to different ways of handling multiple loans for each borrower. Each regression includes branch and day-of-week FE. Standard errors are clustered at the branch-day level. }
%Table \ref{multiple_loans} shows our results are similar when dealing with the multiplicity of loans per borrower.

%\textit{Do file: } \texttt{decomposition\_main\_te.do}
\end{table}

%\hfill \newline
%\noindent \textbf{Results.} 
\paragraph*{Results} The results are stark. The Forced Commitment arm yields large and significant decreases in the cost of loans to clients, as measured either by financial cost or APR. Despite causing an increase in fees, the Forced arm leads to a decrease of 204 pesos in the costs of borrowing (out of a group mean of 942 in the status quo), equivalent to 22\% reduction as a fraction of mean cost. These cost savings arise from a 6.6 percentage point (pp) decrease in the probability of default (out of a baseline mean of 44pp,  implying cost savings of 79 pesos), and also from lower interest payments since, as we will document below, the commitment contract speeds up payments so the interest rate applies to a smaller principal. This translates into a large reduction in APR. A credit product that has an effective average APR of 57\% in the status quo arm (inclusive of default) is reduced to a cost of 46\% through the imposition of a more regularized payment structure. This is in stark contrast to \cite{Pande} for instance.

In contrast, to the impressive effectiveness of the Forced commitment arm, the Choice arm fails to deliver significant changes in any measure, with the exception of an increase in fees, for which we are highly powered, since this outcome is zero for every control observation. Giving borrowers the choice of contract did not decrease financial cost, whereas forcing them into a structured payment contract dramatically reduced it. As we explore later in the paper, the null effect of the Choice arm arises because few borrowers demanded commitment (consistent with \cite{Ashraf}, \cite{Gine}, \cite{Ted}, \cite{Royer}, \cite{Sprenger}), with 89\% choosing the less effective status-quo contract.

%\hfill \newline
%\noindent \textbf{Intermediate outcomes.} 
\paragraph*{Payment speed} To shed light on the mechanisms behind the ATEs discussed above, Table \ref{mechanisms} shows how commitment affects a number of intermediate outcomes. One can group the types of intermediate outcomes into two categories: measures of the speed of pawn recovery, and measures of the decision of when to default. While the first payment for borrowers in the status-quo contract occurs on average only on day 82 (on a 90 days contract), borrowers in the forced commitment arm start paying 13.8 days earlier on average (col 1). Not only do they start paying earlier, the first payment is also 9.7\% larger (col 2), and a larger fraction (7.9\%)  actually pay in full and recover their pawn in the first payment, compared to 30\% in the status quo contract (col 3). The resolution of the loan (either by payment or default) is shortened by 27.9 days (col 4), and conditional on recovery (an endogenous control) by 17.9 days (col 5).

\paragraph*{Payment Bifurcation} A very undesirable outcome from the borrower's perspective is to pay towards the loan without paying in full, i.e. defaulting on the loan while still making some payments. In this case, they lose both the collateral and any payments made toward recovery. One could be concerned that by encouraging them to pay monthly, more borrowers might end up in this dire scenario in the Forced commitment contract. Column 6 shows this is not the case. On the contrary, 7 percentage points fewer borrowers end up in this situation, compared to 12 percent in the status quo contract. Conditional on defaulting those assigned to the Forced commitment contract have paid 4.1\% less of their loan (col 7), and a 14 percent higher proportion of borrowers pay zero conditional on defaulting, an outcome analogous to
``selling their pawn'' (col 8). One interpretation of this bifurcation is that the Forced commitment contract forces borrowers to think earlier about whether they will indeed be able to eventually recover their pawn, and separates borrowers into those ``selling'' their pawn and those recovering it, reducing the share of undecided borrowers that end up paying interest and end up losing the pawn anyway. This mechanism may also help to explain why the Forced commitment contract does not increase the number of visits to the branch to pay (col 10): those recovering their pawn visit more, but those defaulting have 0.20 fewer visits (col 11). Finally, Column 9 shows that treatment effects are concentrated in the intensive margin as treatment does not affect the fraction of clients who pay a positive amount towards pawn recovery.

%\noindent \textbf{Other costs.} 
\paragraph*{Other Costs} Although the paper focuses on financial costs, we consider three additional costs here. First, we include the cost of visiting the branch to make a payment. This includes the self-reported transport cost (most use public transport), as well as the opportunity cost of time. To err on the conservative side, we subtract a whole day's minimum wage the day of the visit, instead of just the wage corresponding to a couple of hours. Second, we consider a rough proxy of the value of liquidity that borrowers lose by paying sooner. To do this we add the interest costs on to the payments in the forced commitment arm and recompute treatment effects with these payments compounding daily (as if they had to borrow in order to make the more rapid payments). Thirdly, so far we have valued the collateral at the gold value appraised by the lender, but the piece may be worth more to the borrower than its gold value.  For many of them the pawned jewelry has sentimental value. This is reflected in the subjective valuation they reported in the survey which is 87\% higher on average. Our third extra cost considers this higher value. Table \ref{table_robustness_fc} shows results are robust to all these changes.

We want to be very clear that we are not measuring borrower welfare, for several obvious reasons. For example, requiring monthly payments may well increase stress we are not taking into account. We are not taking into account discounting given the short period, although in Section \ref{why_paternalism}, we show that to rationalize the findings, we would need unreasonably high discount rates. We are also not measuring financial cost on other loan products.\footnote{However, we think it is unlikely that borrowers ask for even more expensive loans to pay the monthly payment. The `punishment' from the monthly fee is small, less than the transport cost to go to the branch. It is likely better to incur the fee than go through the process of asking for an additional loan to avoid this fee.} Having said this, a large literature focuses on the cost of servicing loans. We follow in the steps of this literature.

%\hfill \newline
%\noindent \textbf{Repeat Pawns.}
%Another way to push the analysis towards a consideration of borrower welfare is the conjecture that if clients exposed to the installment contract like it, they will be more likely to come back vis-a-vis those that got status quo contract.\footnote{\cite{Laibson2018} hypothesizes that consumers use ``experienced utility'' as a guide to make future purchases in this way.} This is indeed what we find in the data.  We consider this outcome in Table \ref{repeat_loans}. We can first ask whether the individual ever takes a loan after the initial study loan, an analysis which is simply experimental.

%\hfill \newline
%\noindent \textbf{Censoring of Loan Completion.}
\paragraph*{Censoring of Loan Completion}
The window of time during which we were able to observe borrower behavior was limited in each branch, meaning that there were loans that we do not see completed (particularly those pawns that were rolled over for one or two further 90-day spells).  Overall, 13\% of all experimental loans are censored, meaning that they neither default nor repay within the observation window.  In the prior analysis we handle this issue in a conservative way by using outcomes such as ``did not default'' which are well-defined even when we do not observe the completion of the loan, resulting in an estimate biased toward zero by the more rapid loan repayment observed in the Forcing arm.  In the Online Appendix we consider this issue in more detail.  Most importantly, Table \ref{bounding_censoring} conducts a bounding exercise that examines how large the effects of this censoring could possibly be by making bracketing assumptions about repayment on censored loans in the treatment and control, respectively.  Comfortingly, Panel B of this table shows that even the most muted possible effect in the bounding exercise still recovers impacts of Forcing on financial costs and APR that are negative and significant at the 99\% level.  Using a lasso-logit model to predict the outcome of censored loans, Panel E shows the APR impact of Forcing increases from a 14 pp reduction (headline results) to a 17 pp reduction.  Hence there appears to be no scope for this censoring issue to overturn our results, and our core results (implicitly assuming censored loans are paid off) is almost certainly an under-estimate of the true impacts.

\paragraph{Lender profit} One of the questions we are interested in is whether lender profits decrease with structured payment. If that is the case, lenders will have little incentive to include this feature in their contracts. For a single loan, lenders get what borrowers lose in a zero-sum way so that a borrower's Financial Cost$_i$ for loan $i$ is the profit from that loan for the lender. Table \ref{main_impact_table} showed that \textit{per loan} borrowers pay less to lenders under the structured payments contract. However, in Table \ref{repeat_loans} in the appendix, we show that borrowers assigned to the forced arm are 6.7\% more likely than in the status quo arm to be repeat customers during our sample period. This brings some extra benefit to lenders that we need to take into account in a simple back-of-the-envelope calculation. From column 1 of Table \ref{main_impact_table} we know \{Financial Cost$_i$ $|$ Contract=c\} for $c$= status quo or structured payment contracts are respectively \hl{XX} and \hl{XX}. From Table \ref{repeat_loans} we know $\Pr(\text{Repeat} | \text{Contract}=c)$ for $c$= status quo or structured payment contracts are respectively \hl{XX} and \hl{XX}. We can use these four numbers to calculate a discounted sum of profits for client $j$ comparing a scenario where only status quo loans are offered with a second scenario where only structured payment loans are offered.\footnote{We do this for simplicity. Otherwise, we would have to assume something about how borrowers switch dynamically across loans, and this is something we do not observe in the data.} One way to do a back-of-the-envelope calculation is by plugging those numbers in the following equation: 
 $\text{Profit}_{j,c} = \{\text{Financial Cost}_j | \text{Contract}=c\}\{1+n\times\Pr(\text{Repeat}_j)\}$ where $n$ is are exogenously arising need to borrower, which are lead to repeat pawning. A second possibility is to compound the probability of coming back to take into account that pawning may lead to more repeat pawning and this compounds: $\text{Profit}_{j,c}  = \sum_{T}\delta^T\{\text{Financial Cost}_j | \text{Contract}=c\}\times\Pr(\text{Repeat} | \text{Contract}=c)^T$, where $\delta$ is the discount rate. We find in the first case that the status quo contract dominates the structured payment contract for all $n$. \hl{For the second alternative, we find that for every  $\delta$ in [0,1] and for every $T$ the status quo also dominates.} That is, the extra repeat probability in the structured payment is not enough to compensate for the loss per each of those loans. Table \hl{XX} in the appendix shows the calculation for different values of $\delta, n, T$.

% \begin{equation}
% \label{profit_function}
% \text{Profit}_{j,c}  = \sum_{T=0}^{\infty}\{\text{Financial Cost}_j | \text{Contract}=c\}\times\Pr(\text{Repeat Pawn}_j | \text{Contract}=c)^T 
% \end{equation}

% where $j$ indexes a borrower, and $c$ indicates the first assigned contract/treatment to borrower $j$. The profit for the status-quo contract is  \$1578, while that of the forced-commitment contract is \$1418, which results in less profit for the forced-commitment contract even when considering repeat pawning behavior.

\section{Choice and Heterogeneous Treatment Effects}
\label{Choice}

The results from Section \ref{Experiment} show that commitment \emph{works}: clients who were assigned to the forced commitment arm experienced substantially lower financial costs on average.
In spite of this, given the opportunity, only 11\% of borrowers chose commitment. 
If the effect of commitment were homogeneous, this would be enough to conclude that the 89\% who did not choose it would have been financially better-off if they had.
In a world of heterogeneous treatment effects, however, low demand for commitment could still be consistent with borrowers adhering to a standard model of rational choice. 
The borrowers who did not choose commitment could simply be those who don't need it. 
Indeed, we find strong evidence that the effect of forced commitment varies substantially across individuals in our experiment: we test and reject
%Appendix \ref{append:chernozhukov} tests and rejects 
the null hypothesis of homogeneous treatment effects using the methodology of \cite{chernozhukov2018generic} (details available upon request).
So the question remains: can we explain the 89\% of borrowers not choosing commitment as being those that don't benefit financially from it?  
In this section, importantly, we show that commitment would lower average financial costs even for the subset of borrowers who choose \emph{not} to commit voluntarily.
To simplify the exposition in this and all sections that follow, we re-define all outcome variables so that \emph{beneficial} treatment effects are \emph{positive}. Using this convention, a positive treatment effect of commitment on financial cost, for example, reflects the average cost \emph{savings} caused by commitment.

\subsection{Bounding the Distribution of Individual Treatment Effects}
\label{sec:bounds}

Calculating the number of individuals who benefit from universal forced commitment requires the distribution of individual treatment effects.
While this distribution cannot be point identified, it can be bounded.
Let $(Y_{i0}, Y_{i1})$ be $i$'s potential outcomes under the control and forced commitment conditions and define $\Delta_i \equiv Y_{i0} - Y_{i1}$.
Let $(F_0, F_1, F_\Delta)$ be the marginal distributions of $(Y_{i0}, Y_{i1}, \Delta)$ and define 
\[
\underline{F}(\delta) \equiv \max \left\{0, \sup_y F_1(y) - F_0(y - \delta)  \right\}, \quad
\overline{F}(\delta) \equiv 1 + \min \left\{0, \inf_y F_1(y) - F_0(y-\delta) \right\}.
\]
Since $F_0$ and $F_1$ are point identified under random assignment, so are $\underline{F}$ and $\overline{F}$, and the pointwise sharp bounds for $F_\Delta$ are $\underline{F}(\delta) \leq F_\Delta(\delta) \leq \overline{F}(\delta)$ \citep{fan2010sharp}.
Figure \ref{fan_park_bounds} plots these bounds for the APR outcome in our experiment.  
To bound the share of borrowers who benefit from forced commitment, we merely substitute $\delta=0$ into the preceding, since $\mathbb{P}(\Delta_i > 0) = 1 - F_\Delta(0)$.
Our point estimates of $\underline{F}(0)$ and $\overline{F}(0)$ are 0.03 and 0.77 respectively, with 95\% confidence intervals of [0.025, 0.050] and [0.75, 0.80]. Since we are interested in $\mathbbm{P}(\Delta_i >0) = 1 - F_\Delta(0)$, it follows that \textit{at least 23\%}, and at most 97\%, of individuals borrowers benefit from forced commitment.%\footnote{Confidence intervals are constructed using the asymptotic distribution of $(\underline{F}, \overline{F})$. See \cite{fan2010sharp}.} 
That is, under general conditions, these bounds already provide evidence for depressed demand, in the sense that at least 12\% (=23\% benefit -11\% demand) of borrowers would incur a lower financial cost with structured payment and yet did not chose it. The bounds are wide however. By adding assumptions it is possible to say more. We turn to this next. %For example, under rank invariance\footnote{i.e.\ if $i$'s rank in the distribution of $Y_0$ equals her rank in the distribution of $Y_1$} the distribution of treatment effects is point identified and given by $F_\Delta(\delta) = \int_0^1 \mathbbm{1}\{ F_1^{-1}(u) - F_0^{-1}(u)\leq \delta\}\,\mathrm{d}u$, where $F_1^{-1}$ and $F_0^{-1}$ are the quantile functions of $Y_1$ and $Y_0$.\footnote{Figure \ref{te_rankinvariance} in Appendix \ref{bounds_FOSD} plots our estimates of $F_\Delta$ under rank invariance.} Because in our data, $Y_1$ first-order stochastically dominates $Y_0$ for financial benefit ($F^{-1}_1(u) - F^{-1}_0(u) >0$), we conclude that all borrowers would benefit ($\Delta_i > 0$).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Original Version before trimming for ECTA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%If we knew the distribution of individual treatment effects, it would be easy assess how many individuals would win and lose under a policy of universal forced commitment, along with the magnitude of any individual harm. 
%Because we can never simultaneously observe the same borrower in both the control and forced commitment arms, however, the distribution of individual treatment effects cannot be point identified.
%It can, however, be bounded.
%We begin our exploration of heterogeneous treatment effects by computing assumption-free bounds on the share of individuals who benefit from forced commitment. 
%
%Let $Y_{i0}$ be borrower $i$'s potential outcome under the control condition, $Y_{i1}$ be her potential outcome under forced commitment, and $\Delta_i \equiv Y_{i0} - Y_{i1}$.
%Because it randomly assigns borrowers to the control and forced arms, our experimental design point identifies the marginal distributions of $Y_{i0}$ and $Y_{i1}$, call them $F_0$ and $F_1$.
%Now, define the functions $\underline{F}$ and $\overline{F}$ as follows:
%\[
%\underline{F}(\delta) \equiv \max \left\{0, \sup_y F_1(y) - F_0(y - \delta)  \right\}, \quad
%\overline{F}(\delta) \equiv 1 + \min \left\{0, \inf_y F_1(y) - F_0(y-\delta) \right\}.
%\]
%Since $F_0$ and $F_1$ are point identified, so are $\underline{F}$ and $\overline{F}$.
%\cite{fan2010sharp} show that the sharp (best possible) pointwise bounds for $F_\Delta$ are given by $\underline{F}(\delta) \leq F_\Delta(\delta) \leq \overline{F}(\delta)$.
%These bounds are simple to compute, and can be surprisingly informative. 
%Here we use them to bound the fraction of borrowers who benefit from forced commitment. 
%Given the way that we have defined our outcome variables, this is the share of borrowers whose treatment effect is positive, i.e.\ $\mathbb{P}(\Delta_i > 0) = 1 - F_\Delta(0)$. 
%To construct the sharp bounds for this quantity, we simply substitute $\delta = 0$ into the preceding equations and estimate $F_0$ and $F_1$ using their empirical analogues constructed from the forced commitment and forced no-commitment arms of the experiment.
%
%Figure \ref{fan_park_bounds} plots the sharp bounds for $F_{\Delta}$ based on the APR outcome. Our point estimates of $\underline{F}(0)$ and $\overline{F}(0)$  are 0.03 and 0.77 respectively, with associated 95\% confidence intervals of [0.025, 0.050] and [0.75, 0.80]. Since we are interested in $\mathbbm{P}(\Delta_i >0) = 1 - F_\Delta(0)$, this means that \textit{at least 23\%} of individual borrowers benefit from commitment (and at most 97\%).\footnote{Confidence intervals are constructed using the asymptotic distribution for the bounds. See \cite{fan2010sharp} for details.}
%
%The bounds we have just described are constructed by considering all possible joint distributions of $Y_0$ and $Y_1$ that are compatible with the observed marginal distributions $F_1$ and $F_0$. 
%With stronger assumptions, it is possible to say more.
%One such assumption is \emph{rank invariance}, which posits that a person's rank in the distribution of $Y_0$ equals her rank in the distribution of $Y_1$.
%While strong, this assumption or variants of it have appeared in a number of settings in the literature, e.g.\ \cite{chernozhukov2005iv}, so we consider it here for purposes of comparison.
%Under rank invariance, the distribution of treatment effects is point identified and given by 
%\[
%F_\Delta(\delta) = \int_0^1 \mathbbm{1}\{ F_1^{-1}(u) - F_0^{-1}(u)\leq \delta\}\,\mathrm{d}u 
%\]
%where $F_1^{-1}$ and $F_0^{-1}$ are the quantile functions of $Y_1$ and $Y_0$. Figure \ref{te_rankinvariance} in Appendix \ref{bounds_FOSD} plots our estimates of $F_\Delta$ under rank invariance.
%For the financial benefit outcome, $Y_1$ first-order stochastically dominates $Y_0$ in our experiment. Since this implies that $F^{-1}_1(u) - F^{-1}_0(u)$ is always positive, rank invariance yields an extremely stark conclusion: all borrowers have positive individual treatment effects of forced commitment. 
%For the APR outcome, the results under rank invariance are slightly less stark: we estimate that just under 90\% of borrowers have positive individual treatment effects. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% End of original version before trimming for ECTA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Potential Outcomes and Exclusion}
\label{sec:potentialOutcomes}

%The assumption-free \cite{fan2010sharp} bounds from the preceding section show that more than 23\% of borrowers would benefit from a policy of forced commitment. This suggests that some of the 89\% of borrowers in the choice arm who did \emph{not} choose commitment would have faced lower borrowing costs if they had.Making this intuition precise, however, requires a careful consideration of the relationship between choice and heterogeneous treatment effects. 
This section shows under an additional mild exclusion restriction assumption, our 3-arm RCT design point-identifies a range of treatment effects of interest, and in particular, the treatment on the untreated, necessary to study the benefits mandating a contractual feature. To this end, we now provide a full definition of the potential outcomes in our empirical setting. Let $Z_i \in \{0, 1, 2\}$ denote the treatment arm to which to participant $i$ was assigned: $Z_i = 0$ denotes the forced no-commitment arm, $Z_i = 1$ denotes the forced commitment arm, and $Z_i = 2$ denotes the choice arm. 
Now let $D_i$ be the treatment that participant $i$ actually \emph{received}, where $D_i = 0$ denotes no-commitment and $D_i = 1$ denotes commitment. 
We assume perfect compliance in the $Z_i = 0$ and $Z_i = 1$ arms. %\footnote{For more discussion on this point, see Section \ref{sec:integrity} above.}
It is only in the $Z_i = 2$ arm that participants are free to choose between alternative contracts. 
Let $C_i \in \{0, 1 \}$ denote a participant's ``choice type.'' If $C_i = 1$ then participant $i$ \emph{would choose commitment}, given the option; if $C_i = 0$ she would not. 
As shorthand, we call borrowers with $C_i = 1$ ``choosers'' and those with $C_i = 0$ ``non-choosers.''
Whereas a participant's choice type $C_i$ is only observed if she is allocated to the choice arm ($Z_i = 2$), her treatment $D_i$ and experimental arm $Z_i$ are always observed. 
Given the design of our experiment, these quantities are related by
\begin{equation}
D_i = \mathbbm{1}(Z_i \neq 2) Z_i + \mathbbm{1}(Z_i = 2) C_i.
\label{eq:potentialTreatments}
\end{equation}

We maintain the stable unit treatment value assumption (SUTVA) throughout. 
This means that borrower $i$'s outcomes depend only on her \emph{own} values of $Z_i$ and $D_i$, not those of any other person in the experiment. 
Under this assumption, a fully general model for the potential outcomes in our experiment would take the form $Y_i(d, z)$ for $d\in \{0,1\}$ and $z \in \{0, 1, 2\}$, allowing participant $i$'s potential outcome to depend \emph{both} on the treatment she actually receives, $D_i$, and the experimental arm to which she is assigned, $Z_i$. 
This model is too general, however, to point identify meaningful causal effects. 
For this reason, we consider two mild exclusion restrictions.

To state these restrictions we first define some additional notation.
Because our experimental design implies that any borrower with with $Z_i = 0$ has $D_i = 0$, we abbreviate the potential outcome $Y_i(d=0,z=0)$ as $Y_{i0}$. 
Similarly, since any borrower with $Z_i = 1$ has $D_i = 1$, we abbreviate $Y_i(d=1,z=1)$ as $Y_{i1}$. 
%This is in keeping with our notation from section \ref{sec:bounds} above.
Using this notation, our first exclusion restriction, which we call the ``No-choosing warm-glow'' (NCW) assumption is given by
\begin{equation}
Y_i(d=0,z=2) = Y_i(d=0,z=0) \equiv Y_{i0}.
\label{eq:exclusion0}
\end{equation}
Equation \ref{eq:exclusion0} only restricts the potential outcomes of non-choosers,  individuals with $C_i = 0$, because they are the only borrowers for whom $D_i = 0$ when $Z_i = 2$.
In words, this condition assumes that every non-chooser experiences the same potential outcome regardless of whether she is assigned to the choice arm or the control arm.
Similarly, we assume a second NCW exclusion restriction is given by
\begin{equation}
Y_i(d=1,z=2) = Y_i(d=1,z=1)\equiv Y_{i1}.
\label{eq:exclusion1}
\end{equation}
Equation \ref{eq:exclusion1} only restricts the potential outcomes of choosers, individuals with $C_i =1$, because they are the only borrowers for whom $D_i = 1$ when $Z_i = 2$.
In words, this condition assumes that every chooser experiences the same potential outcome regardless of whether she is assigned to the treatment arm or the choice arm.

Mathematically \eqref{eq:exclusion0} and \eqref{eq:exclusion1} have the same structure as the standard LATE exclusion restriction that $Y_i(d,z)$ depends only on $d$, not on $z$. 
Substantively, however, they are slightly different, given that there is no explicit reference to the ``chosen'' versus ``forced'' treatment distinction in the usual LATE setup.
In essence, \eqref{eq:exclusion0} and \eqref{eq:exclusion1} assume that being assigned a particular treatment has the same result as choosing it for yourself, provided that you are assigned the same treatment that you \emph{would have chosen}. 
If the mere fact of having been given a choice has a direct effect on outcomes, one or both of our exclusion restrictions will be violated.
One can imagine situations in which this might be the case.
For example, even someone who would have voluntarily chosen to undergo drug rehabilitation, given the choice, might respond differently when coerced.
In our empirical setting, however, both \eqref{eq:exclusion0} and \eqref{eq:exclusion1} are plausible.\footnote{Closely related assumptions are common, if often tacit. \cite{chamberlain2011bayesian} explicitly assumes that choosing and being assigned a treatment have the same effect. A significant literature using compulsory schooling laws to estimate the returns to education tacitly assumes that schooling \emph{in general} has the same returns, regardless of whether it is mandated or chosen freely. Similarly, fertilizer yields are tacitly assumed to generate the same returns regardless of whether the fertilizer was chosen by farmers or provided by the government, as long as it is the same fertilizer.} 
Moreover, each has testable implications that we fail to reject: see \autoref{sec:testing_exclusion}. 
Under \eqref{eq:exclusion0} and \eqref{eq:exclusion1}, the observed outcome $Y_i$ is related to $(Y_{i0}, Y_{i1})$ by 
\begin{equation}
    Y_i = \mathbbm{1}(Z_i =0) Y_{i0} + \mathbbm{1}(Z_i = 1)  Y_{i1}  + \mathbbm{1}(Z_i = 2) \left[(1 - C_i) Y_{i0} + C_i Y_{i1} \right].
\label{eq:potentialOutcomes}
\end{equation}
Equation \ref{eq:potentialOutcomes} is the key to understanding the results that follows. 
Random assignment of $Z_i=0$ and $Z_i = 1$ identifies the marginal distributions of $Y_{i0}$ and $Y_{i1}$ for the population as a whole. 
Random assignment of $Z_i=2$ likewise point identifies the share of choosers ($C_i = 1$), the distribution of $Y_{i1}$ for choosers, and the distribution of $Y_{i0}$ for non-choosers ($C_i = 0$).
Because $Z_i$ is assigned independently of pre-treatment covariates $X_i$, we also identify the \emph{conditional} distributions of $Y_{i0}$ and $Y_{i1}$ given $X_i$. 
%These distributions are the ingredients that we use below to explore the importance of paternalism.

%The fact that the Forced Commitment contract lowers financial costs for almost all clients, together with the fact that only 11\% choose it must mean that many consumers are choosing contracts with higher financial costs for themselves. We now use several different techniques to shed further light on the idea that the `wrong' people are choosing commitment.

%Our core impacts show that forced commitment is uniquely effective, a result that can only arise if those who would not have freely chosen commitment benefit from it.  We now delve more deeply into this result.

%\subsection{Treatment Effect Heterogeneity}

%Defenders of choice make another subtler point: that even if take up of the commitment-fee contract is low, the sole fact of \textit{having} choice could potentially change behavior or welfare (\cite{Dalboetal:2010}, \cite{Sjostrometal:2018}, \cite{Tlaxcala}).

%In spite of broadly distributed benefits of the Forced Commitment commitment contract we documented above, 90\% of clients chose \textit{not} to have it, preferring the status-quo contract. As explained above we are fairly confident this was not because a lack of understanding of the contracts.

%\vspace{.1in}
%\noindent What can explain this low demand? In the spirit of \cite{Blumenstock} we try to differentiate between explanations. One explanation could be risk. Even if the fee forcing contract generates cost savings on average, clients with risk averse preferences may not demand it if they perceive higher risk from it.   Alternatively, even risk-neutral agents may not demand it if the cost of lost flexibility is high or if they have high temporal discount rates and prefer to pay later. %If clients, for instance, face large and frequent income shocks which make it hard to repay monthly.\footnote{We don't observe income shocks in our data, making is hard to quantify how important they are. 91\% of those in the Forced Commitment arm incurred a fee, which suggests that shocks are not uncommon.} 
%Although the evidence that follows undermines several explanations, the tests we can conduct given the data are not sharp enough to leave a single explanation standing. Section \ref{model} lays our a simple theory (namely present bias with partial naivete) which fits all the 6 main findings of the paper. Given its parsimony and explanatory power, we submit that it is the most suitable explanation for our results.

%Following two  paragraphs moved from old 'behavioral' choice section towards end of paper
%\hfill \newline
%\subsection{The effect of choice on pawn recovery and financial cost} \label{effect_choice}

%Figure \ref{main_te} represents the  effect sizes (in standard deviations) of the two main study arms for three outcomes:  default, the financial cost of the loan, and the effective interest rate paid to acquire the borrowed capital.  The Forced arm is significant for all three outcomes, with very substantial treatment effects in magnitudes for default and APR, and the commitment arm is never significant.\footnote{The slight increase in APR for the Choice arm seen in this picture arises from the fact that the APR calculation has the arm-specific loan size in the denominator, and the loans in the Choice arm turn out to be slightly smaller (although statistically equal) on average than the Control.}    In other words, voluntary commitment is ineffective but the use of forced commitment saves the average borrower an amount of money almost twice the APR of a normal US credit card loan.

%Although we have shown that Forced Commitment reduces financial cost and increases pawn recovery, this does not imply that welfare increases. Clients could have disutility from the stress of having monthly installments for instance. 
%We don't need to make the strong claim of welfare for the purposes of this paper. The financing cost results are striking enough. But we

%The panel labeled Forced Commitment in Figure \ref{reincidence} presents results of estimating equation (\ref{basic_reg}) with the dependent variable being a dummy for the client being a repeat customer. We define a client as a repeat customer if, after experiencing at least 75 days of the respective treatment arm, she came back and pawned a \textit{different} piece.\footnote{Results are robust to using number of days experiencing the contract larger than 75. We know it is a different piece from it having a different weight/value.}  The leftmost coefficient shows that the causal effect of the Forced Commitment contract on repeat pawning is an increase of 5 percentage points in the probability of being a repeat customer, this is twice the repeat purchasing happening in the status quo group. This effect is not explained by clients recovering their first pawn with higher probability in the Forced Commitment contract and re-pawning this same one since we make sure it is another piece; besides only 16\% recover their piece in the fist 75 days in the fee forcing contract. Experience with the frequent payment contract seems to be necessary as there is no effect of the monthly payment contract on repeat pawning in the first 30 or 60 days of the contract (see Figure \ref{reincidence_before} in the Appendix).

%\hl{The second coefficient from the left focuses on the subsample who are predicted would have a treatment effect on financial cost savings above the median} treatment effect in their first pawn.\footnote{Since this is predicted using \cite{atheygrf} based on pre-determined covariates for treatment and control clients it is valid to split the sample this way.} It shows that for these clients the likelihood of coming back is twice as large. That is, those that benefit more from the Forced Commitment contract are more likely to repeat compare to their control group. The remaining coefficients of the ``Forced Commitment'' panel condition on endogenous outcomes and should be interpreted with care. The \hl{third} coefficient from the left restricts the sample to those that had not recovered their pawn in the first 75 days in both groups and finds similar results. The \hl{fourth} coefficient conditions on clients who recovered their pawn of the experiment, again on in both arms. Results from these two endogenous samples are similar to those for the whole sample.

%The fourth coefficient \footnote{The second and third estimate must be interpreted with care as we are conditioning on a variable that may be affected by treatment. We view these 2 coefficients as suggestive correlations only. The fourth coefficient does not have that problem as it conditions on clients that by their pre-determined covariates are predicted to have above the median treatment effects (in both arms).}

%The coefficient on repeat pawning is just as big when we focus only 2nd pawns happening after the first was recovered (third coefficient ``fnr'').  This is consistent with \cite{Laibson2018}'s conjecture that clients could be making decisions based on experienced utility.\footnote{The third coefficient conditions on the subsample that did recover their first pawn in both arms and finds a coefficient of 4pp. This is not a causal estimate as recovery is influenced by treatment, but we found it telling that the differences were similar to the causal estimate for the full sample.}

\subsection{The ``Forcing-Choice'' Design}
\label{sec:randchoice}

%Above we used outcome data from the forced arms ($Z_i\neq 2$) along with the commitment take-up rate in the choice arm ($Z_i = 2$) to argue that at least some of the \hl{89\%} of participants who did \emph{not} choose commitment would have been better off if they had. 
%By additionally examining \emph{outcome data} in the choice arm, it is possible to say more. 
As a direct consequence of \eqref{eq:potentialOutcomes}, our experimental design---henceforth the ``forcing-choice design''---point identifies a number of interesting and economically-relevant causal quantities. 
First it identifies the treatment on the treated (TOT) and untreated (TUT) effects:  
\[
\text{TOT} \equiv \mathbbm{E}(Y_{i1} - Y_{i0} | C_i = 1), \quad
\text{TUT} \equiv \mathbbm{E}(Y_{i1} - Y_{i0} | C_i = 0).
\]
The TOT is the causal effect of commitment on borrowers who would voluntarily choose it, while the TUT is the causal effect on borrowers who would not.
If the TUT is positive, then borrowers who did not choose commitment would have experienced better outcomes, \emph{on average}, if they had. Selection on gains implies TOT$>$TUT$>$ATE, as it is typically assumed in canonical Roy models. Because our design identifies all three quantities, it allows us to test this implication directly and to calculate the average selection on gains (ASG), namely the difference between the TOT and TUT effects:
\[
\text{ASG} \equiv \mathbbm{E}[Y_{i1} - Y_{i0}|C_i = 1] - \mathbb{E}[Y_{i1} - Y_{i0} | C_i = 0] = \text{TOT} - \text{TUT}.
\]

A number of recent papers compare estimates of the TOT and TUT to better understand who selects into treatment and why, e.g.\ \cite{cornelissen2018benefits} and \cite{Walters}. 
This line of work relies, at least to some extent, upon structural modeling assumptions to extrapolate from the reduced-form quantities that are identified by the data alone to more interesting, and economically relevant, causal parameters.\footnote{While the marginal treatment effects (MTE) approach \citep{heckman2007econometric} can in principle be used to identify the TOT and TUT without parametric restrictions, doing so requires an instrumental variable $Z$ with sufficiently rich support that the probability of treatment take-up given $Z$ varies continuously between zero and one. In practice, instrumental variables are usually discrete and, even when continuous, typically have a more modest effect on take-up.} 
An alternative approach aims to avoid structural assumptions by calculating conditional local average treatment effects (LATE) given observed covariates $X$ and re-weighting them according to the distribution of covariates in some population of interest to yield, for example, an average treatment effect \citep{aronow2013beyond,angrist2013extrapolate}. 
%For example, one might re-weight using the distribution of $X$ in the population as a whole, rather than the sub-population of compliers, to extrapolate from LATE to ATE. 
But there is no free lunch: this ``LATE-and-reweight'' approach relies upon assumptions of its own, most crucially the assumption that there is \emph{no selection-on-gains} conditional on $X$, i.e.\ that the conditional LATE equals the conditional ATE. 
In contrast to both approaches, the forcing-choice design uses exogenous experimental variation to point identify the ATE, TOT, and TUT without ruling out unobserved selection-on-gains or relying on additional structural modeling assumptions.

\begin{figure}
\caption{Graphical Intuition for the Forcing-Choice Design.} 
    \begin{center}
        \centering
        \includegraphics[width=1.0\textwidth]{Figuras/tot_tut_intuition.png}
    \end{center}
 \scriptsize{
     %This figure illustrates why the controlled choice design identifies the TOT and TUT effects.
    Gray rectangles denote borrowers with a commitment contract; white rectangles denote borrowers with a status quo contract. 
    A comparison of means across control and forcing arms identifies the ATE of forcing commitment. 
    %In the choice arm, anyone with a commitment contract is a chooser; in the control and forcing arms, we do not know who is a chooser and who is not, but because borrowers are randomly assigned to experimental treatment arms, the share of choosers will be the same, on average (depicted using dashed vertical lines). 
    The difference of mean outcomes across the choice and control ``nets out'' the non-choosers, and hence equals the TOT multiplied by the share of choosers.
    %arms gives the intent-to-treat effect of offering choice. 
    %Under the exclusion restriction that moving non-choosers between the choice and forcing arms does not change their outcomes, this comparison ``nets out'' the non-choosers.
    %Hence, the ITT of offering choice equals the TOT multiplied by the share of choosers. 
    %Similarly, under the exclusion restriction that moving choosers between the choice and forcing arm does not affect their outcomes. 
    Similarly, the difference of means across the forcing and choice arms ``nets out'' the choosers and equals the TUT multiplied by the share of non-choosers.
        The share of choosers, illustrated using dashed vertical lines, is equal on average across arms under random assignment.}
    \label{tot_tut_graph}
     %\textit{Do file: }  \texttt{tot\_tut\_graph.do}
\end{figure}

The key insight can be read directly from \eqref{eq:potentialTreatments} and \eqref{eq:potentialOutcomes}; Figure \ref{tot_tut_graph} provides graphical intuition. %for our identification approach. % derivations appear in \ref{append:randchoice}
Viewing $Z_i$ as an instrumental variable, the forcing-choice design can be interpreted as a \emph{pair} of RCTs, each subject to one-sided non-compliance.
The first of these compares $Z_i=0$ to $Z_i = 2$. For each individual with $Z_i = 0$ we have $D_i = 0$ and observe $Y_{i0}$. For those with $Z_i = 2$ we have $D_i = C_i$ and observe $(1 - C_i) Y_{i0} + C_i Y_{i1}$. This is identical to a ``randomized encouragement'' design in which treatment is only available to those who are encouraged: $Z_i = 2$. Under this interpretation, those with $C_i = 1$ are ``the compliers'' and it follows by a standard argument (see Section \ref{sec:proofs}) that 
\begin{equation}
\frac{\mathbbm{E}(Y_i|Z_i=2) - \mathbbm{E}(Y_i|Z_i =0)}{\mathbbm{E}(D_i|Z_i=2)-\mathbbm{E}(D_i|Z_i=0)} = 
\frac{\mathbbm{E}(Y_i|Z_i=2) - \mathbbm{E}(Y_i|Z_i =0)}{\mathbbm{E}(D_i|Z_i=2)} = \text{TOT} 
\label{eq:TOT}
\end{equation}
since $\mathbbm{E}(D_i|Z_i=0)=0$ by \eqref{eq:potentialTreatments}. 
A closely related argument can be used to construct a Wald estimand that identifies the TUT. Here we consider $Z_i = 1$ to be the ``encouragement'' and compare the outcomes for these individuals to those with $Z_i = 2$. If $Z_i = 1$ then $D_i = 1$ and we observe $Y_{i1}$.
If instead $Z_i = 2$ then $D_i = C_i$ and we observe $(1 - C_i) Y_{i0} + C_i Y_{i1}$. Again, we can view this as an experiment with one-sided non-compliance, but now the situation is reversed. Everyone with $Z_i = 1$ is treated, but some people with $Z_i = 2$ are ``always-takers'' who obtain the treatment ($D_i = 1$) despite having been allocated to the ``control'' arm $Z_i=2$. Under this interpretation, the ``compliers'' are those with $C_i = 0$: when $Z_i=1$ they take the treatment, and when $Z_i=2$, they do not. Thus, 
\begin{equation}
\frac{\mathbbm{E}(Y_i|Z_i=1) - \mathbbm{E}(Y_i|Z_i =2)}{\mathbbm{E}(D_i|Z_i=1)-\mathbbm{E}(D_i|Z_i=2)} = 
\frac{\mathbbm{E}(Y_i|Z_i=1) - \mathbbm{E}(Y_i|Z_i =2)}{1 - \mathbbm{E}(D_i|Z_i=2)} = \text{TUT}
\label{eq:TUT}
\end{equation}
since $\mathbbm{E}(D_i|Z_i = 1)= 1$ by \eqref{eq:potentialTreatments}. 
Because they identify both the TOT and TUT, \eqref{eq:TOT} and \eqref{eq:TUT} also identify the average selection on gains: $\text{ASG} = \text{TOT} - \text{TUT}$.

The forcing-choice design also identifies the average selection bias (ASB) and average selection on levels (ASL). In particular,
\begin{align}
\label{eq:ASB}
    \text{ASB} &\equiv \mathbbm{E}(Y_{i0}|C_i=1) - \mathbbm{E}(Y_{i0}|C_i = 0) = \frac{\mathbbm{E}(Y|Z=0) - \mathbbm{E}(Y|Z=2,D=0)}{\mathbbm{E}(D|Z=2)}\\
    \label{eq:ASL}
    \text{ASL} &\equiv \mathbbm{E}(Y_{i1}|C_i = 1) - \mathbbm{E}(Y_{i1}|C_i=0) = \frac{\mathbbm{E}(Y|Z=2,D=1) - \mathbbm{E}(Y|Z=1)}{1 - \mathbbm{E}(D|Z=2)}
\end{align}
as shown in Section \ref{sec:proofs}.
The ASB tells us whether borrowers who voluntarily choose commitment are those who are worse off, on average, under the status quo.
Similarly, the ASL tells us whether borrowers who voluntarily choose commitment are those who are better off, on average, under the commitment contract.

Equations \eqref{eq:TOT}--\eqref{eq:ASL} are useful for understanding why the forcing-choice design identifies the TOT, TUT, ASG, ASB, and ASL but they are less convenient for estimation and inference. 
\autoref{sec:estimation_inference} explains how to compute each of these quantities from a small  number of just-identified, linear instrumental variables regressions, along with appropriate cluster-robust standard errors.
These estimators and standard errors are implemented in our companion STATA package. %\textcolor{red}{Add a link to this package along with the documentation.}

%\todo[inline]{Possibly add some discussion here of Craig's idea: using the controlled choice design to study ``targeting''}

%A unique feature of our design is the randomized assignment to treatment (forced commitment), control (forced no-commitment), and choice of treatment.

%This structure allows us to observe the average value of both the treated and untreated potential outcomes, as well as the selection into treatment that would occur under choice.  A starting point for how to exploit this structure is to apply the logic usually used to estimate the Treatment on the Treated (ToT) to also recover the Treatment on the Untreated (TUT).  This approach assumes nothing about the decision to choose (other than random sampling generating counterfactual compliance rates that would have been the same in every arm), but imposes three exclusion restriction-style assumptions typical of Local Average Treatment Effect (LATE) analysis.  Namely, a) the typical assumption that an individual not choosing the treatment has the same potential outcome as one not offered it, then b) the mirror image assumption for the Forcing arm that an individual forced to take Commitment has the same potential outcome as an individual freely choosing it, and c) no spillovers from chooser to non-choosers in the choice arm.\footnote{\hl{For robustness we also try a Manski-type monotone IV approach with similar results.}}

%More formally, let $Z_i$ be the observed, randomly assigned experimental allocation, where $Z_i=0$ denotes the control arm, $Z_i=1$ denotes the Forced arm, and $Z_i=2$ is the Choice arm.  Let $C_i$ be the choice type; $C_i=0$ when individual $i$ does not choose commitment and $C_i=1$ when chooses commitment, observed in the Choice arm and latent in the other two arms.  A fraction $p$ of individuals chooses the treatment in the Choice arm.  Finally, $D_i = Z_i\times \mathds{1}(Z_i\neq 2) + C_i\times \mathds{1}(Z_i=2)$ is the observed treatment indicator, and $Y(d,z)$ is the potential outcome function for $d=0,1$, and $z=0,1,2$.

%To test the difference between gains for choosers versus gains from non-choosers we need to identify $\mathbb{E}[Y_1-Y_0\;|\;C=1]$ (the Treatment on the Treated), and $\mathbb{E}[Y_1-Y_0\;|\;C=0]$ (the Treatment on the Untreated).   The observed outcome in the Choice arm is the weighted average of the treated and untreated outcomes; $p \times\mathbb{E}[Y_1\;|\;C=1] + (1-p)\times\mathbb{E}[Y_0\;|\;C=0]$.  Given assumptions a) and c) above we can recover the ToT from the difference between the Choice and Control arms as is standard, and then symmetrically we can invoke b) and c) to recover the TUT from the difference between the Forced and Choice arms as follows:

%\begin{align*}
%ToT \equiv \mathbb{E}[Y_1 - Y_0 \;\mid\; C=1]  &=   \frac{\mathbb{E}[Y \;\mid\; Z=2] - \mathbb{E}[Y \;\mid\; Z=0]}{p} \\
%TuT \equiv \mathbb{E}[Y_1 - Y_0 \;\mid \;C=0]  &=  \frac{\mathbb{E}[Y \;\mid\; Z=1] - \mathbb{E}[Y \;\mid\; Z=2]}{(1-p)} 
%\end{align*}

%This estimate of the ToT is the standard one that would result from instrumenting for compliance with the offering of treatment as in \cite{angrist1996identification}.  The TuT is the analogous estimator, which could be recovered in a regression pooling the forced and choice arms and instrumenting for \textit{not} complying with being assigned to the choice arm (for a more formal development, see the appendix).  The advantage of representing the estimands via division of the relevant ITT by the compliance (non-compliance) rate is that it allows us to estimate all the necessary terms from the single pooled regression Equation \ref{basic_reg}, from which the significance levels for the ToT, the TuT, and the difference between them can all be estimated as single-equation F-tests:

%\[ToT = \frac{\beta^C}{p}; \qquad\quad TuT = \frac{(\beta^F - \beta^C)}{(1-p)}; \qquad\quad (ToT-TuT) = \frac{\beta^C}{p}- \frac{(\beta^F - \beta^C)}{(1-p)}. \]

%\hfill \newline
%\noindent \textbf{TOT and TUT results.}

Table \ref{tot_tut} calculates the causal quantities described above---TOT, TUT, ASG, ASB, and ASL---for our experimental data, along with robust standard errors for each. 
For purposes of comparison, the table also presents the ATE results from Section \ref{Experiment} above (row 1),\footnote{Coefficients are not exactly the same since Table 4 includes branch and day-of-week FE.} along with the corresponding average potential outcomes $\mathbbm{E}[Y_0]$ and $\mathbbm{E}[Y_1]$ (rows 4--5).
The columns of the table correspond to different outcome variables defined above.
We are now in a position to report one of the important results in the paper. For all four outcome definitions, the TUT effect is positive, statistically and economically significant, and comparable in magnitude to the ATE.
In other words: the structured payments contract is on average \emph{beneficial} to the people who \emph{would not choose it}, and these benefits are large. That is, mandating structured payment contracts would decrease the financial cost. Borrowers are leaving money in the table by not choosing it.

Some comments are in order. First, this result holds on average. Some non-choosers could lose from a mandate. Section \ref{sec:RF} assesses personalized targeting strategies. Second, financial cost is not welfare, and further work with more comprehensive measures (stress, consumption, performance in other loans) needs to be done. We provide proof of concept that at least financial cost may not be minimized by free choice and provide a method to rigorously compare take-up mandates against consumer choice. Third, given that providing such contracts may not be in the lender's interest, borrowers may not even be presented with this choice.

Table \ref{tot_tut} also estimates TOT. Due to the low take-up rate of commitment in the choice arm, TOT effects are imprecisely estimated. Only one of them, \%(1-Default) from column (3), is statistically significant. 
This carries over into our estimates of the average selection on gains. Our point estimates are \emph{negative} for all but the (1 - Default) outcome, but none is statistically distinguishable from zero.  
%Although we cannot quite reject the null hypothesis that the TUT and TOT effects are equal--our power is limited because of the low take-up rate of commitment in the choice arm--the pattern of coefficients is consistent with some degree of selection-on-gains.
%Commitment is more beneficial to people who are willing to choose it than it is to people who are not.
For the (1 - Default) outcome, we have sufficient precision to conclude that the average selection bias (ASB) is large and negative.
This means that borrowers who choose commitment would have faced a higher probability of default under the status quo contract than borrowers who do not choose commitment.%\footnote{We may not want to read too much into TOT vs TUT comparisons as they are imprecise. But taken at face value, the result that TOT$>$TUT for default, while the opposite is true for financial cost, suggests that, while voluntary commitment has a slightly stronger effect on allowing borrowers to avoid default, forcing \textit{reduces} payments in a strong enough manner as to overcome the default-benefits of choice.}

Overall, Table \ref{tot_tut} suggests that commitment works but that \emph{not enough} people choose to commit: the commitment contract is beneficial on average even to those who would \emph{not} choose it voluntarily. We believe this result is new in the household finance literature. It also illustrates how the ``Forcing-Choice'' design can be used to study more generally whether low take-up of an intervention is problematic, based on the impacts among non-choosers.\footnote{For instance, in the debate on financial commitment take-up, some papers argue it is low (\cite{Ashraf}, \cite{Gine}, \cite{Ted}, \cite{Royer} while others argue it is high (\cite{Kremer}, \cite{Casaburi}, \cite{Alcohol}, \cite{AprajitP&P}, \cite{Pascaline}), but none estimates the benefits for non-takers. In a different domain, doctors claim that medical treatment abandonment is too high in a broad range of diseases \citep{non_adherence}, without knowing the treatment benefits for those that abandon.}

\begin{table}[H]
\caption{Five Treatment Effects Estimates: TOT, TUT, ASG, ASB, ASL}
\label{tot_tut}
\begin{center}
\footnotesize{\input{./Tables/tot_tut.tex}}
\end{center}
\scriptsize{This table presents results computed using the derivations from Section \ref{sec:randchoice}. The APR and financial cost outcomes have been multiplied by $-1$ so that a positive causal effect \emph{benefits} the borrower in each of the four columns. The bottom panel presents p-values for a number of null hypothesis tests of treatment effect heterogeneity. }
%\textit{Do file: } \texttt{tot\_tut.do}
\end{table}

%Below we explore possible explanations for this result.
%and the ``right people'' choose to commit: those who are most likely to benefit from it and those whose outcomes are most adverse under the status quo.

%Given that the ITT results for the Choice arm imply an insignificant worsening of outcomes relative to the control, our estimates of the Treatment on the Treated in the first row of the bottom panel show negative results that are larger in absolute magnitude (multiplied by ~9, the inverse of the compliance rate) and similarly insignificant.  The Treatment on the Untreated (TuT), conversely, is strongly positive across the board, and indicates a significant improvement for all outcomes (here ITTs are effectively multiplied by 1.12, the inverse of the non-compliance rate).  Most novel is the ability to form a test of the difference between the ToT and the TuT, which is conducted in the bottom rows of the table.  We provide four different ways of estimating p-values for this comparison; the standard clustered estimates motivated by the experimental design, normal and percentile bootstraps, and randomization inference.  The differences between treatment effects for compliers and non-compliers are of borderline significance in some of the unadjusted specifications, but are not significant once adjusted.\footnote{It is worth noting that the statistical power of these estimands is driven by the compliance rate; given our low overall compliance we are better powered to measure the TuT than the ToT, and the power for the difference between the two would be highest with a compliance rate of 50\%.}   Because of the large standard errors on the ToT coming from the low compliance rate in the Choice arm we are unable to state with statistical confidence that the TuT is higher than the ToT, but the TuT is strongly positive while the ToT is weakly negative, and so in both relative and absolute terms the choices being made are `wrong' relative to the potential outcomes.

\section{The Case for Paternalism}
\label{Paternalism}

\subsection{Why does paternalism work in this context?}
\label{why_paternalism}

A behavioral literature has highlighted voluntary commitment as an attractive way of allowing the ``right'' people to self-select. % A discussion of compulsory commitment is necessarily paternalistic, and by its nature focuses on impacts among those who would not have selected commitment voluntarily. 
%Viewed through the lens of rational choice theory, it is unsurprising that we estimate $\text{TOT}>\text{TUT}$. 
The argument for compulsory treatment, centers on the surprising result that $\text{TUT}>0$. We now investigate four potential explanations for the positive $\text{TUT}$: the need to learn, time discounting, present bias, and overconfidence.

%\hfill \newline
%\noindent \textbf{Learning.}

Our experiment introduced a new contract into an environment that had not previously featured commitment; perhaps clients required experience to understand its benefits. Are clients who experienced the commitment contract more likely to choose commitment subsequently than those assigned to the status-quo contract? %We have already shown above that those who experienced the Forcing contract were more likely to borrow again.  We now ask whether the subset who were assigned to the Choice arm when they returned were more likely to choose commitment.  
Appendix Table \ref{learning_table} %\footnote{Table \ref{learning_table} presents information about participants' \emph{immediate subsequent} pawning behavior.  For borrowers who returned to pawn again more than once, this analysis considers only their first repeat pawn.}
shows \textit{no difference} in choosing commitment for those assigned to forced commitment versus those assigned to status quo control that came back to pawn again (22\% did). %The implication is that while those who have experienced commitment feel more positively towards the pawn contract, the experience does not lead them to conclude that they need commitment on the subsequent loan.  
While these exercises cannot completely exclude the possibility that learning would play a role with more exposure, they provide no indication that the lack of voluntary compliance is simply a matter of inexperience with commitment in the short run. Note also that there will be no learning if lenders don't provide a structured payment contract in the first place.

%Column (1) considers the 228 clients who returned only a second time to pawn again at a day/branch that was randomly assigned to the choice arm. Each of the two rows in this column presents a difference of mean commitment take-up rates, and associated standard error. The first row compares those who were \emph{initially} assigned to forced commitment against those where were assigned to control; the second row compares those who were initially assigned to the choice commitment arm to those who were assigned to the other two arms. In each case, there is no statistically discernible difference in the rates of commitment take-up. Granted, this is a selected sample because the decision to pawn again is potentially endogenous to the initial treatment allocation. For this reason, Column (2) considers the full sample of 4441 borrowers by re-defining the outcome variable to be an indicator for returning to pawn again at a branch/day when commitment was offered \emph{and} choosing commitment. This composite outcome variable is not subject to the sample selection problem (although it is directly driven by the decision to repeat borrow). The comparison in the two rows remains the same: forced commitment versus control in row one and choice commitment versus forced arms in row two. Again, there is no statistically discernible difference in commitment take-up rates in either row. While these exercises cannot completely exclude the possibility that learning plays a role, they provide no indication that the lack of voluntary compliance is simply a matter of inexperience with commitment.

%\hfill \newline
%\noindent \textbf{Time discounting.}

Can high borrower time discount rates explain low demand for structured payments? %Highly impatient individuals might rationally choose the \emph{status quo} contract, despite the benefits that commitment yields returns, 
After all, monthly payments are front-loaded while pawn recovery is back-loaded, even if by only a few days.  
%those returns are realized at contract ending hile the cost of monthly payments are incurred  Our estimated decrease in the financial cost of credit from above ignores discounting, but commitment involves incurring up-front costs (early payment) in return for delayed benefits (a higher probability of recovering one's pawn). 
Figure \ref{fc_discount_rates} calculates TUT effects for different borrower discount rates given the actual timing of repayment and pawn recovery. The solid line gives the TUT effect adjusted for a specified annual discount rate, while the shaded regions give the associated 95\% \& 90\% confidence interval.
We see that non-choosers continue to experience significant decreases in the net present value of financial costs up to annual discount rates of 1,000\%, and the NPV remains positive, although not significant only at 5,000\% discount rates. These are huge discount rates.\footnote{In a similar exercise for subprime borrowers in the US, \cite{Levin} argue that even discount rates smaller than these would be too large to be reasonable and can be safely discarded as explanations.} As such, discounting is unlikely to explain why those who benefit, on average, from commitment fail to choose it. % when offered.

%\hfill \newline
%\noindent \textbf{Present bias.} 
If the benefits of commitment among non-choosers cannot be explained by standard models of rational choice, the canonical behavioral story would center on time inconsistency.  While commitment is useful to anyone with hyperbolic time preferences, only those who are sophisticated---i.e.\ aware that they are hyperbolic discounters---will demand it.  A large share of ``na\"ive'' hyperbolics in the population--borrowers who are unaware that they are hyperbolic discounters--could therefore drive a large and positive $\text{TUT}$.  Our baseline survey included standard questions about discount rates between today and a month in the future versus discount rates between three and four months in the future.
This allows us to classify borrowers who display more impatience over immediate delays as present biased. This measure of financial hyperbolicity is widely used in survey research, although it is not without problems.\footnote{Our measure is dichotomous, and it is not incentivized. Recent empirical work has shown the superiority of more elaborate measures such as ``convex time budgets'' \citep{andreoni2015measuring} while questioning the interpretation of measures of hyperbolicity that are not based on consumption \citep{andreoni2012estimating, cohen2020measuring}, suggesting that real effort tasks provide a better measure \citep{augenblick2015working}.  Given that we had only a few minutes to interview real pawnshop clients prior to a commercial transaction, our simple measure was a necessary compromise.}

If we could perfectly measure present bias and sophistication, we could divide the sample into three groups: sophisticated hyperbolics (who chose commitment), time-consistent non-choosers (for whom forcing will not be effective), and na\"{i}ve hyperbolic non-choosers (who will benefit from forced commitment). %If present bias fully explains the low take-up rate of voluntary commitment, we should find that the TUT for present-biased borrowers  is positive while the TUT for all other borrowers is not.
If present bias fully explains the low take-up rate of voluntary commitment, we should find that the TUT for present-biased borrowers is positive. This is because among the group of non-takers, a comparison of present-biased borrowers against everyone else is a comparison of na\"{i}ve hyperbolics against time-consistent non-choosers. The left panel of Figure \ref{tut_beh_partition} in the Appendix carries out a feasible version of this exercise using our survey measure of present bias.
%The overall TUT estimate along with a 95\% confidence interval is given in blue.\footnote{For all borrowers who answered our present-bias survey questions.}
%The corresponding TUT estimate and confidence interval for present-biased borrowers identified with the survey question is given in green; results for all other borrowers are shown in red.
%The overall TUT is a weighted average of the impact in these two sub-groups.  The TUT among the present biased is insignificant and less than half the size of the strongly significant TUT among those who are \textit{not} present biased. Therefore, taking our survey measure of hyperbolicity at face value. 
We find no indication that present-bias explains our positive estimated TUT.

%\hfill \newline
%\noindent \textbf{Sure confidence.}

Finally, motivated by the high subjective probabilities of recovery we found in the baseline survey, and by the fact that a large fraction of borrowers pay towards recovery and then lose the pawn and these payments, we explore overconfidence in recovery as a potential explanation of low demand for structured payments. After all, recall that while 72\% of survey respondents believe they have a 100\% chance of recovering their pawn, in reality only 43\% will go on to do so.  
Borrowers who are certain to recover their pawn perceive no benefit of committing to repay under the threat of fees and, therefore, no benefit of choosing the structured payment contract.
%if individuals who \emph{believe} that they are certain to repay, choose (rationally given their incorrect beliefs) to forgo the costs associated with commitment that are designed to induce repayment. 
To explore this, as in the preceding paragraph, we compare the overall TUT estimate to estimates computed for two sub-groups, defined by a binary variable that we call ``sure-confidence''. This measure equals one for any individual who says at the time of borrowing that they have a 100\% probability of recovering their pawn, and zero otherwise.
%The right panel of Figure \ref{tut_beh_partition} presents the results for this exercise.
%The overall TUT estimate for all borrowers who answered our sure-confidence survey questions is given in blue, along with a 95\% confidence interval.
%The corresponding TUT estimate and confidence interval for sure-confident borrowers is given in green; results for all other borrowers are shown in red.
In contrast to our results for present bias, we find that sure-confidence \textit{is} a strong predictor of TUT. Figure \ref{tut_beh_partition} shows that the TUT is almost entirely confined to the sure confident individuals, while those that say they have some chance of defaulting have TUT close to zero. It seems that the effectiveness of paternalism in our experiment may be driven by \emph{overconfident} borrowers who, heedless of the risk of default, fail to choose commitment despite benefiting substantially when they are forced to commit.\footnote{In Figure \ref{determinants_sure} we plot the coefficient estimates from a regression that predicts sure confidence with a battery of individual-level characteristics.  Older males are more likely to be sure-confident, as are those with more education.  Taken at face value, the sure-confident also report less financial stress, less trouble paying bills, and to be more frequently relied upon financially by family members.  Viewed through a behavioral lens, however, it is also possible that the type of person who is over-confident in their ability to repay a loan also exaggerate their degree of economic security in their response to survey questions. In any case, it appears that sure confidence may be difficult to predict with easily-observed and objective demographic criteria, a point to which we return below.}

\subsection{Targeting Paternalism using Causal Forests} \label{sec:RF}

%\noindent \textbf{Causal forests.} 
On average, the commitment contract benefits both those who would choose it and those who would not.
In Section \ref{sec:bounds}, we briefly went \emph{beyond} average effects by presenting bounds on the distribution of individual treatment effects.
Because they made no assumptions beyond random assignment, these bounds were relatively wide.
%Adding the assumption of rank invariance yielded a distribution of treatment effects in approximately 90\% of borrowers had positive individual treatment effects for the APR outcome, implying that \emph{practically everyone} would benefit from treatment and hence that there would be hardly any no losers from paternalism. Rank invariance, however, is an extremely strong assumption.\footnote{For the financial cost outcome, rank invariance implies that all invididual treatment effects are positive.}
We then applied NCW assumptions to estimate average treatment effects for choosers and non-choosers.

In this section, we again impose NCW exclusions, but this time conditional on observed covariates, not just on average. This enables us to use causal forests to estimate \textit{conditional} average treatment effects (CATE) as well as CTOT and CTUT effects. This stronger assumption allows us to say more about personalized treatment effects and, therefore, to assess different ways to target structured payments.  %This exercise provides more fine-grained information about treatment effect heterogeneity.
Among other things, it will potentially allow us to identify groups of borrowers who are on average \emph{harmed} by commitment.
%Under the stronger assumption that our observed survey measures capture the main sources of treatment effect heterogeneity, this exercise will allow us to approximate individual-level counterfactuals, to consider whether particular borrowers made ``mistakes'' in their choice of contract. 
To estimate the conditional average treatment effects we use the ``generalized random forest'' methods of \cite{atheygrf} (see Appendix \ref{app:cate} for details).\footnote{The generalized random forest approach of \cite{atheygrf} produces conditional average effect estimators that are asymptotically normal.
Our inferences in this section are carried out by simulating from the normal limit distributions using the estimated standard errors. As covariates, we use 22 questions in our survey (what we call the \textit{wide} set of covariates in Appendix \ref{app:cate}. The survey questionnaire (with the highlighted questions being the ones for the wide set of covariates) can be found in \ref{baseline_survey} in the Appendix.} 
We implement these estimates using the covariates from our survey.

\vspace{.2in}
\begin{figure}[!h]
     \caption{Heterogeneous Treatment Effects.} 
     \label{heterogeneous_effects}    
    \begin{center}
     \begin{subfigure}{0.35\textwidth}
       \centering
      \includegraphics[width=\textwidth]{Figuras/he_dist_tau_hat_tot.pdf}
          \caption{ToT}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
       \centering
      \includegraphics[width=\textwidth]{Figuras/he_dist_tau_hat_tut.pdf}
          \caption{TuT}
    \end{subfigure} 
       \begin{subfigure}{.35\textwidth}
        \centering
        \includegraphics[width=\textwidth]{Figuras/he_dist_tau_hat_eff.pdf}
        \caption{ATE}
    \end{subfigure} 
    \end{center}
  \scriptsize{To estimate conditional average treatment effects given administrative and survey data and the function \texttt{causal\_forest()} of the \texttt{grf} R package; to estimate conditional TOT and TUT effects we use the \texttt{instrumental\_forest()} function from the same package. For more information see Appendix \ref{app:cate}.} 
  %\textit{Do file: }  \texttt{cate_dist.do}
\end{figure}

%\hfill \newline
%\noindent \textbf{Treatment effect distributions.} 
Figure \ref{heterogeneous_effects} plots densities of the estimated CATE, CTOT, and CTUT effects from the generalized random forest models described above.
In each case, the outcome variable is APR benefit, i.e.\ the reduction in APR from a structured payments contract.
As we see from the figure, the conditional average effects are highly heterogeneous but overwhelmingly positive. 
The TUT density is particularly interesting for the question of paternalism since, as emphasized above, it presents conditional average effects for borrowers who would not voluntarily choose commitment.
Only 7\% of our estimated conditional TUTs are negative (95\% confidence interval of [4\%, 9\%]).\footnote{To be clear, this is a probability statement about conditional average effects over the distribution of covariates.
In particular, we estimate that  $\int \mathbbm{1}\{\mathbb{E}[Y_1 - Y_0| X = \mathbf{x}, C = 0] < 0\} \, f(\mathbf{x}|C=0) \, \mathrm{d}\mathbf{x}$ is approximately 0.07. The share of non-choosers with negative conditional \emph{average} treatment effects need not equal the share with a negative \emph{individual} effects, i.e.\ $\mathbbm{P}(Y_1 < Y_0| C = 0)$.
But the more treatment effect heterogeneity that $X_i$ explains, the closer these two values become.}
Figure \ref{heterogeneous_effects} strengthens the argument that many borrowers will save on financial cost by choosing the structured payment contract.

%\hfill \newline
%\noindent \textbf{``Mistake'' thresholds.} 
%Under the assumption that our instrumental forest estimates capture important sources of treatment effect heterogeneity, we can use them to assess whether particular borrowers in the choice arm made ``mistakes'' in their decision to accept or refuse the commitment contract, in terms of predicted financial costs of the loan.
We now use the CTUT estimates to quantify how much extra financial cost would borrowers incur under different scenarios. To simplify language, we define a ``mistake'' for a non-chooser to be a conditional TUT estimate that significantly exceeds some large and positive APR threshold, e.g., \ 10 percentage points. 
At any such threshold, we can calculate the percentage of non-choosers in the choice arm who have benefited by more than that threshold from having chosen commitment. Again, we do not measure welfare, only financial cost.

The results of this exercise can be read off from the green curve in Figure \ref{choose_wrong}.
Defining $F_{\text{TUT}}(\delta)$ to be the CDF corresponding to the density of  CTUT estimates from Figure \ref{heterogeneous_effects}, the green curve in Figure \ref{choose_wrong} is merely $[1 - F_{\text{TUT}}(\delta)] \times 100\%$.
In other words, the vertical axis for the green curve plots the percentage of non-choosers who made a ``mistake'' of a certain threshold from the horizontal axis. Note that we use a \emph{positive} APR threshold to denote a mistake for both choosers and non-choosers.  This ensures that bigger mistakes are always to the \emph{right} of smaller mistakes. 
The green shaded region gives associated 95\% pointwise confidence bounds.

\vspace{.2in}
\begin{figure}[H]
 \caption{``Mistakes'' in the choice arm.}
\begin{center}
        \centering
        \includegraphics[width=0.65\textwidth]{Figuras/line_cw_apr_tot_tut.pdf}
 \end{center}       
 \scriptsize{This figure presents conditional average TUT and TOT effects for the APR outcome from Figure \ref{heterogeneous_effects} in an alternative manner, to consider the fraction of borrowers in the choice arm who made ``mistakes'' in their decision to accept or refuse the commitment contract. A ``mistake'' for a non-chooser is defined as a positive conditional TUT effect that significantly exceeds a specified threshold APR value.
 %\textit{Do file: } \texttt{choose_wrong_quant_wrong_tot_tut.do}

%  The green curve equals $[1 - F_{\text{TUT}}(\delta)]\times 100\%$, where $F_{\text{TUT}}$ is the CDF corresponding to density of conditional TUT effects from Figure \ref{heterogeneous_effects}, computed using the instrumental forest approach from \cite{atheygrf}. Evaluated at any positive value on the horizontal axis, it gives the fraction of non-choosers in the choice arm who made a ``mistake'' by not choosing commitment. The green shaded region gives associated 95\% pointwise confidence bands. 
% Analogously, a ``mistake'' for a chooser is defined as a negative conditional TOT effect that exceeds a specified threshold APR value. The red curve equals $F_{\text{TOT}}(-\delta)$ where $F_{\text{TOT}}$ is the CDF corresponding to the density of conditional TOT effects from Figure \ref{heterogeneous_effects}. 
% The red shaded region gives associated 95\% pointwise confidence bands.
% For both the red and green curves, we define the APR threshold so that larger mistakes are to the \emph{right} of smaller mistakes.
% This allows us to construct the overall fraction of mistakes in the choice arm, the blue curve, as a weighted average of the green and red curves.
% The weights equal the share of choosers and non-choosers in the choice arm.
}
    \label{choose_wrong}
\end{figure}

For choosers we follow an analogous approach, defining a ``mistake'' as a \emph{negative} CTOT exceeding a particular APR threshold.
The results for choosers can be read from the red curve in Figure \ref{choose_wrong}.
If $F_{\text{TOT}}(\delta)$ denotes the CDF corresponding to the density of conditional TOT estimates from Figure \ref{heterogeneous_effects}, then the red curve is merely $F_{\text{TOT}}(-\delta) \times 100\%$.
%In other words, the red curve gives the percentage of choosers who made a ``mistake'' when mistakes are defined at a particular APR threshold.
The red shaded region gives associated 95\% pointwise confidence bounds.
The blue curve in Figure \ref{choose_wrong} gives the \emph{overall} percentage of borrowers in the choice arm who made a ``mistake'' at a particular APR threshold.
This total is computed by taking a weighted average of the green (non-choosers) and red (choosers) curves, with weights equal to their shares in the choice arm.%\footnote{The blue curve is very similar to the green curve because 89\% of borrowers in the choice arm are non-choosers.}

The results in Figure \ref{choose_wrong} suggest that a large fraction of non-choosers would have experienced lower financial costs had they chosen the structured payment contract, even at an APR threshold as large as 10 percentage points, we estimate that more than half of them should have chosen the structured payment contract to lower financial costs.
In contrast, relatively few choosers appear to have made mistakes by choosing it.
This now allows us to make a stronger statement in favor of paternalism in our context; not only does forced commitment generate large benefits \textit{on average}, but it also benefits the \emph{vast majority} of borrowers who would be coerced under a policy of forced commitment.

\subsection{Can we target paternalism accurately?}

Even though the overwhelming majority of borrowers would save on financing costs, not all of them would, according to our estimates. Is it possible to target paternalism only towards those who would benefit from it?

The answer depends on how predictive the covariate information is of treatment effects. Note that to implement a targeting protocol, the covariate information used in the model would need to be either verifiable or hard to manipulate. In our survey and administrative data, this leaves us with only a few covariates: age, gender, high school education or above, desired loan size, and whether that individual has ever pawned before. 
We call these the ``narrow'' covariate set, to contrast with the full set of survey variables, which we call the ``wide'' covariate set.  We take the estimated   $\widehat{\text{CATE}}(X_i)$ using the wide covariates as our ground truth.\footnote{We assign a dummy variable equal to one for the 90\% of borrowers who have a positive CATE using the wide covariate set. Figure \ref{wide_narrow_forests} shows the relationship between these two different  CATEs.  We generate substantially less heterogeneity when using the narrow covariate set, although we still reject the null hypothesis of no treatment effect heterogeneity with the test from \cite{chernozhukov2018generic}.}

To investigate targeting we ask how well we can predict this ground truth using the narrow covariate set, which is the set of less-manipulable variables that the pawnlender has. We use both a logistic regression and a random forest classification model using the narrow covariate set to predict positive benefits at the individual level.  We compare the in-sample performance of targeting rules against a policy of universal forced commitment.

\begin{table}[H]
\caption{Type I \& II errors using targeting narrow rules}
\label{hit_miss_rule}
\begin{center}
\footnotesize{\input{./Tables/hit_miss_rule.tex}}
\end{center}
\scriptsize{This table reports error rates for six different rules for allocating individuals to commitment. Row 1 assigns all borrowers to control, Row 2 all to structured payments. Row 3 uses `optimal targeting according to the CATE from the wide covariate set. Row 4 uses a random forest classification, and Row 5 a logit model, both with only the narrow covariate set.  Row 6 uses the choices made by borrowers. %DoFile : wide_narrow_forests.do
}
\end{table}

Table \ref{hit_miss_rule} shows error rates for six possible assignment rules: assigning all borrowers to control, all to structured payments, the optimal (wide-fores), the random forest with narrow covariates, the Logistic model with narrow covariates, and the actual choice made by borrowers.  All models take the wide RF as the ground truth. While the narrow RF correctly assigns to control roughly half of those who do not benefit from the structures payment contract, it also incorrectly allocates to control 4.38\%  that would gain from structured payments. As such, it only improves the overall correct targeting rate by about half of a percentage point relative to universal Forcing. The Logit assignment rule is even less successful at predicting benefits and harms, with a higher share of borrowers incorrectly assigned to both treatment and control, meaning that the overall correct targeting rate for the Logit is 5 pp \textit{lower} than universal forcing to structured payments.  Self-targeting through choice proves to be little better than assigning everyone to the control condition, given the low take-up rate and the presence of both Type I and Type II errors in the choice arm.

The takeaway is that given (a) low take-up, (b) the large fraction of the sample benefiting from commitment, (c) and the weak predictive power of the narrow covariates,universal paternalism assigning all borrowers to the commitment contract appears to be an attractive targeting method in our context.

\section{Conclusion} \label{conclusion}

This paper makes several contributions. First, it analyzes the important but understudied industry of pawn loans. Likely, hundreds of millions of people use pawn loans, and yet economists have not studied them. We show that a simple change to contract terms results in substantial financial savings for pawn borrowers: forced commitment lowers the APR from 57\% to 46\%, and reduces the fraction of borrowers who default by 6.6 pp, or 15\%. That this new contract generated large benefits for borrowers and yet was not offered, and that a contract that generated default was the industry standard instead, is related to the idea of ``veiled paternalism'' \cite{Laibson2018}, put on its head, %In ``veiled paternalism'' principals embed forms of commitment into their products but mask this fact from consumers who may need but do not desire commitment, in pawn lending, over-collateralization means that lenders stand to make more money from defaulting borrowers, generating incentives for ``veiled \textit{non}-paternalism,'' 
embedding features that lead to high borrower costs.

%Potentially due to the nature of the borrower pool, 
%In our context voluntary commitment choice does improve borrower outcomes, but forcing commitment did induce significant cost savings for the overall group of clients.

Second, a novel ``forcing-choice'' design paired with NCW assumptions allows us to go beyond ATE results and draw an important set of conclusions about the relationship between take-up and heterogeneous treatment effects, providing a complement to marginal treatment effects methods. %In particular, we simultaneously point-identify the impact of commitment on those who would naturally choose it \emph{and} those who only experience commitment when forced. 
Estimating treatment effects for non-takers is critical in thinking about paternalism, and we show one rigorous way to do it that is new in the literature and distinct from Karlan-Zinman designs.

Third, in terms of our empirical results, we find substantial benefits of treatment for \textit{non-choosers} and no evidence of selection on gains by borrowers who choose commitment. This is the first such result in the household finance literature and a first step in thinking about paternalistically imposing frequent payment contracts. Our results show that over-optimism of pawn recovery is the characteristic most strongly associated with benefiting from the commitment despite not having chosen it, and it maybe an explanation for the low demand for commitment. Given that the rate of voluntary commitment in our sample is only 11\% and given the limited benefits from personalized targeting with our data, in order to achieve widespread benefits in this context compulsory commitment appears to be an attractive policy. With a now well-established toolkit of regular small payments and incentives delivering small default rates in microfinance, regulators may fruitfully investigate the possibility of requiring pawnbrokers to embed features of commitment and regularity into their repayment structures in more consistent ways.

%Why do borrowers leave such substantial returns on the table?

%Our borrower pool overestimates their probability of repayment by more than 50\%, and our positive TUT estimate is largely confined to borrowers who incorrectly believe that they have little chance of defaulting.  %More standard explanations such as discounting, learning, or time inconsistency find little support in our data.

%Using machine learning methods, we find the benefits of commitment close to universal. The benefits of targeting commitment based on characteristics that lenders can observe and participants would truthfully reveal are extremely limited, suggesting that universal commitment is an attractive policy in our empirical setting.

%Where lenders have no incentive to engage in veiled paternalism and customers display inefficiently low demand for it, financial policy regulation may prove an attractive option.  Pawnshops, along with other over-collateralized credit products such as payday lending, exist in an environment where the lender may desire customers to lose their collateral on the loan, hurting especially low-income populations who are its main users. With a now well-established toolkit of regular small payments and incentives delivering small default rates in microfinance, regulators may fruitfully investigate the possibility of requiring pawnbrokers to embed features of commitment and regularity into their repayment structures in more consistent ways.\footnote{If employed at scale in a competitive lending sector this would redistribute welfare from those who would have repaid (whose interest rates must now rise to cover lower returns from collateral seizure) towards those who would only repay in the presence of commitment. In a setting of lender market power however, redistribution from lenders to borrowers could occur.}

There are several questions left for future research. Would results hold using more comprehensive measures of welfare? How fast would borrowers learn about the benefits of structured payments if regulators forced lenders to provide this contractual feature? How would lender competition affect the incentives to provide such a contract? Much needs to be done to improve the lives of the millions of borrowers using panw loans.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%BIBLIOGRAPHY
\begingroup
\setstretch{1.10}
\bibliographystyle{authordate1}
%\bibliographystyle{amsalpha}
%\bibliographystyle{AER}

\bibliography{References}
\endgroup

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Proofs below
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Proofs}
\label{sec:proofs}

This section gives a formal derivation of the identification results presented in Equations \eqref{eq:TOT}--\eqref{eq:ASL} of \autoref{sec:randchoice}.
To simplify the presentation, we omit $i$ subscripts throughout.
%and adopt the shorthand  $Z_0 \equiv \mathbbm{1}(Z=0)$, $Z_1 \equiv \mathbbm{1}(Z=1)$, and $Z_2 \equiv\mathbbm{1}(Z=2)$.
%The following assumption collects our exclusion restriction and the key features of the constrained choice design.

\begin{assumption}[Randomized Choice Design and Exclusion Restriction]\mbox{}
\label{assump:randchoice}
   \begin{enumerate}[(i)]
   \item $Z$ is independent of $(Y_{0}, Y_{1}, C)$
   \item $D = \mathbbm{1}(Z \neq 2)Z + \mathbbm{1}(Z = 2)C$
   \item $Y = \mathbbm{1}(Z = 0) Y_{0} + \mathbbm{1}(Z = 1) Y_{1} + \mathbbm{1}(Z = 2) [ (1 - C) Y_{0} + C Y_{1}]$
   \end{enumerate}
\end{assumption}

\begin{lem}
Under Assumption \ref{assump:randchoice},
\label{lem_randchoice}
   \begin{enumerate}[(i)]
       \item $\mathbbm{E}(D|Z=2) = \mathbb{P}(C=1)$
       \item $\mathbbm{E}(Y|Z=0) = \mathbbm{E}(Y_0)$
       \item $\mathbbm{E}(Y|Z=1) = \mathbbm{E}(Y_1)$
       \item $\mathbbm{E}(Y|D=0,Z=2) = \mathbbm{E}(Y_0|C=0)$
       \item $\mathbbm{E}(Y|D=1,Z=2) = \mathbbm{E}(Y_1|C=1)$.
   \end{enumerate} 
\end{lem}

\begin{proof}
Part (i) follows because $Z=2$ implies $D=C$ and $Z$ is independent of $C$.  
Parts (ii) and (iii) follow similarly: given $Z=0$ we have $Y = Y_0$, given $Z=1$ we have $Y = Y_1$, and $Z$ is independent of $(Y_0,Y_1)$.
For parts (iv) and (v), first note that Assumption \ref{assump:randchoice} (iii) implies that $Z$ is conditionally independent of $(Y_0,Y_1)$ given $C$.
Now, $Z=2$ implies that $D=0$ if and only if $C=0$. Hence, $\mathbbm{E}(Y|D=0, Z=2) = \mathbbm{E}(Y_0|C=0)$ establishing part (iv).
For part (v) $Z=2$ implies that $D=1$ if and only if $C=1$ from which it follows that $\mathbbm{E}(Y|D=1, Z=2) = \mathbbm{E}(Y_1|C=1)$.
\end{proof}

\begin{prop} 
Under Assumption \ref{assump:randchoice},
    \begin{enumerate}[(i)]
        \item $\text{TOT} \equiv \mathbbm{E}(Y_1 - Y_0|C=1)  = \displaystyle \frac{\mathbbm{E}(Y|Z=2) - \mathbbm{E}(Y|Z=0)}{\mathbbm{E}(D|Z=2)}$
        \item $\text{TUT} \equiv \mathbbm{E}(Y_1 - Y_0|C=0) = \displaystyle \frac{\mathbbm{E}(Y|Z=1) - \mathbbm{E}(Y|Z=2)}{1 - \mathbbm{E}(D|Z=2)}$
        \item $\text{ASB} \equiv \mathbbm{E}(Y_0|C=1) - \mathbbm{E}(Y_0|C=0) = \displaystyle \frac{\mathbbm{E}(Y|Z=0) - \mathbbm{E}(Y|Z=2,D=0)}{\mathbbm{E}(D|Z=2)}$
        \item $\text{ASL} \equiv \mathbbm{E}(Y_1|C=1) - \mathbbm{E}(Y_1|C=0) = \displaystyle \frac{\mathbbm{E}(Y|Z=2,D=1) - \mathbbm{E}(Y|Z=1)}{1 - \mathbbm{E}(D|Z=2)}$.
    \end{enumerate}
\end{prop}

\begin{proof}
Parts (i) and (iii) we require an expression for $\mathbbm{E}(Y_0|C=1)$ in terms of $(Y, D, Z)$.
By Lemma \ref{lem_randchoice}(ii) and iterated expectations 
\[
\mathbbm{E}(Y|Z=0) = \mathbbm{E}(Y_0) = \mathbbm{E}(Y_0|C=0) \mathbbm{P}(C=0) + \mathbbm{E}(Y_0|C=1) \mathbbm{P}(C=1).
\]
Re-arranging and substituting Lemma \ref{lem_randchoice}(i) and (iv),
\begin{align}
\mathbbm{E}(Y_0|C=1)  
%&= \frac{\mathbbm{E}(Y|Z=0) - \mathbbm{E}(Y_0|C=0) \mathbbm{P}(C=0)}{\mathbbm{P}(C=1)}\nonumber\\ 
&=  \frac{\mathbbm{E}(Y|Z=0) - \mathbbm{E}(Y|Z=2,D=0) \mathbbm{E}(1 - D|Z=2)}{\mathbbm{E}(D|Z=2)}.
\label{eq:Y0C1}
\end{align}
Part (i) follows by combining \eqref{eq:Y0C1} with Lemma \ref{lem_randchoice}(v) and simplifying; part (iii) follows by combining \eqref{eq:Y0C1} with Lemma \ref{lem_randchoice}(iv) and simplifying.
Similarly, for parts (ii) and (iv) we require an expression for $\mathbbm{E}(Y_1|C=0)$ in terms of observables.
By Lemma \ref{lem_randchoice}(iii) and iterated expectations,
\[
\mathbbm{E}(Y|Z=1) = \mathbbm{E}(Y_1) = \mathbbm{E}(Y_1|C=0)\mathbbm{P}(C=0) + \mathbbm{E}(Y_1|C=1) \mathbbm{P}(C=1).
\]
Re-arranging and substituting Lemma \ref{lem_randchoice}(i) and (v),
\begin{align}
\mathbbm{E}(Y_1|C=0) 
%&= \frac{\mathbbm{E}(Y|Z=1) - \mathbbm{E}(Y_1|C=1)\mathbbm{P}(C=1)}{\mathbb{P}(D=0)} \nonumber\\
&=\frac{\mathbbm{E}(Y|Z=1) - \mathbbm{E}(Y|Z=2,D=1)\mathbbm{E}(D|Z=2)}{\mathbb{E}(1-D|Z=2)}.
\label{eq:Y1C0}
\end{align}
Part (ii) follows by combining \eqref{eq:Y1C0} with Lemma \ref{lem_randchoice}(iv) and simplifying; part (iv) follows by combining \eqref{eq:Y1C0} with Lemma \ref{lem_randchoice}(v) and simplifying.
\end{proof}

\newpage

\input{Restud submission/OA_restud}
\end{document}